---
title: "3-Class-LPA & Predictors"
author: "Simon"
date: 2026-04-16
format: html
editor: visual
echo: true
messages: false
warnings: false
bibliography: references.bib
---

## Preprocessing

```{r warning = FALSE, message = FALSE}
#| label: activate packages
library(readxl)
library(writexl)
library(car)
library(tidyverse)
library(tidyLPA)
library(mclust)
library(ggplot2)
library(ggalluvial)
library(forcats)
library(nnet)
library(broom)
library(knitr)
library(ggeffects)
library(interactions)
```

```{r}
#| label: load data
setwd("C:/Users/simon/Nextcloud2/Shared/LPA - Simon/04_longitudinal_analyses")

mbi_merged = read_xlsx("bl_f01_f02_classes.xlsx")
head(mbi_merged)
```

## Mean differences between profiles

Calculate mean scores per profile and compute mean differences:

```{r}
#| label: mean scores

mean_bl = mbi_merged |> 
  select(case, exhaustion_bl, inefficacy_bl, cynicism_bl, bl_class) |> 
  pivot_longer(
    cols = c(exhaustion_bl, inefficacy_bl, cynicism_bl),
    names_to = "bl_dimension",
    values_to = "bl_score") |> 
  group_by(bl_class, bl_dimension) |> 
  summarise(mean_score_bl = mean(bl_score),
            .groups = "drop")

mean_f01 = mbi_merged |> 
  select(case, exhaustion_f01, inefficacy_f01, cynicism_f01, f01_class) |> 
  pivot_longer(
    cols = c(exhaustion_f01, inefficacy_f01, cynicism_f01),
    names_to = "f01_dimension",
    values_to = "f01_score") |> 
  group_by(f01_class, f01_dimension) |> 
  summarise(mean_score_f01 = mean(f01_score),
            .groups = "drop")

mean_f02 = mbi_merged |> 
  select(case, exhaustion_f02, inefficacy_f02, cynicism_f02, f02_class) |> 
  pivot_longer(
    cols = c(exhaustion_f02, inefficacy_f02, cynicism_f02),
    names_to = "f02_dimension",
    values_to = "f02_score") |> 
  group_by(f02_class, f02_dimension) |> 
  summarise(mean_score_f02 = mean(f02_score),
            .groups = "drop")

mean_scores = cbind(mean_bl, mean_f01, mean_f02)
```

Create a plot with overlaid timepoint profiles:

```{r warning = FALSE, message = FALSE}
#| label: mean superimposed plot
#| echo: false

mean_scores = mean_scores |> 
  mutate(bl_dimension = str_remove(bl_dimension, "_bl$"),
         f01_dimension = str_remove(f01_dimension, "_f01$"),
         f02_dimension = str_remove(f02_dimension, "_f02$"))

mean_scores_long = bind_rows(
  mean_scores |>
    select(bl_class, bl_dimension, mean_score_bl) |>
    rename(class = bl_class, dimension = bl_dimension, mean_score = mean_score_bl) |>
    mutate(timepoint = "bl"),
  mean_scores |>
    select(f01_class, f01_dimension, mean_score_f01) |>
    rename(class = f01_class, dimension = f01_dimension, mean_score = mean_score_f01) |>
    mutate(timepoint = "f01"),
  mean_scores |>
    select(f02_class, f02_dimension, mean_score_f02) |>
    rename(class = f02_class, dimension = f02_dimension, mean_score = mean_score_f02) |>
    mutate(timepoint = "f02")) |>
  mutate(
    dimension = factor(dimension, levels = c("exhaustion", "cynicism", "inefficacy")),
    color_key = paste(class, timepoint, sep = "_"))

# Color vector
profile_colors = c(
  "Burnout_bl" = "midnightblue", "Burnout_f01" = "steelblue3",
  "Burnout_f02"= "lightskyblue", "Disengaged_bl" = "firebrick4",    
  "Disengaged_f01" = "indianred3", "Disengaged_f02" = "palevioletred2",
  "Overextended_bl" = "orangered3", "Overextended_f01" = "orange2",       
  "Overextended_f02" = "lightsalmon", "Ineffective_bl" = "darkslategray", 
  "Ineffective_f01" = "darkolivegreen", "Ineffective_f02" = "darkolivegreen3",
  "Engagement_bl" = "darkgoldenrod3", "Engagement_f01" = "goldenrod2",    
  "Engagement_f02" = "gold")

mean_scores$mean_score_bl = as.numeric(mean_scores$mean_score_bl)
mean_scores$mean_score_f01 = as.numeric(mean_scores$mean_score_f01)
mean_scores$mean_score_f02 = as.numeric(mean_scores$mean_score_f02)

fig1 = ggplot(mean_scores_long,
               aes(x = dimension, y = mean_score,
                   group = color_key, color = color_key)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = profile_colors) +
  scale_x_discrete(labels = c(
    exhaustion = "Exhaustion",
    cynicism   = "Cynicism",
    inefficacy = "Inefficacy"
  )) +
  scale_y_continuous(limits = c(0, 6)) +
  labs(y = "Mean Score", x = "", color = "Profile") +
  theme_minimal()

plot(fig1)
#setwd("C:/Users/simon/Nextcloud2/Shared/LPA - Simon/04_longitudinal_analyses/plots")
#ggsave("mean_scores.png", fig1, width = 18, height = 16, bg = "white", units = "cm", dpi = 400)

```

```{r warning = FALSE, message = FALSE}
#| label: mean differences between BL + Follow-Ups
#| echo: false

# Burnout
burnout_ttest = mbi_merged |>
  filter(bl_class == "Burnout",
         f01_class == "Burnout") |>
  select(case,exhaustion_bl, cynicism_bl, inefficacy_bl,
    exhaustion_f01, cynicism_f01, inefficacy_f01)

# Paired t-tests for each dimension
t_exhaustion = t.test(burnout_ttest$exhaustion_bl, burnout_ttest$exhaustion_f01, paired = TRUE)
t_cynicism   = t.test(burnout_ttest$cynicism_bl,   burnout_ttest$cynicism_f01,   paired = TRUE)
t_inefficacy = t.test(burnout_ttest$inefficacy_bl, burnout_ttest$inefficacy_f01, paired = TRUE)

library(broom)

bind_rows(
  tidy(t_exhaustion) |> mutate(dimension = "exhaustion"),
  tidy(t_cynicism)   |> mutate(dimension = "cynicism"),
  tidy(t_inefficacy) |> mutate(dimension = "inefficacy")) |>
  select(dimension, estimate, statistic, p.value, conf.low, conf.high) |> 
    kable(
    caption = "Paired t-tests: Burnout profile at baseline vs. follow-up 1",
    col.names = c("Dimension", "Mean Difference", "t", "p", "CI low", "CI high"),
    digits = 3)

# Disengaged
disengaged_ttest = mbi_merged |>
  filter(bl_class == "Disengaged",
         f01_class == "Disengaged") |>
  select(case,exhaustion_bl, cynicism_bl, inefficacy_bl,
    exhaustion_f01, cynicism_f01, inefficacy_f01)

# Paired t-tests for each dimension
t_exhaustion = t.test(disengaged_ttest$exhaustion_bl, disengaged_ttest$exhaustion_f01, paired = TRUE)
t_cynicism   = t.test(disengaged_ttest$cynicism_bl,disengaged_ttest$cynicism_f01,   paired = TRUE)
t_inefficacy = t.test(disengaged_ttest$inefficacy_bl, disengaged_ttest$inefficacy_f01, paired = TRUE)

bind_rows(
  tidy(t_exhaustion) |> mutate(dimension = "exhaustion"),
  tidy(t_cynicism)   |> mutate(dimension = "cynicism"),
  tidy(t_inefficacy) |> mutate(dimension = "inefficacy")) |>
  select(dimension, estimate, statistic, p.value, conf.low, conf.high) |> 
    kable(
    caption = "Disengaged profile at baseline vs. follow-up 1",
    col.names = c("Dimension", "Mean Difference", "t", "p", "CI low", "CI high"),
    digits = 3)

# Overextended
overextended_ttest = mbi_merged |>
  filter(bl_class == "Overextended",
         f01_class == "Overextended") |>
  select(case,exhaustion_bl, cynicism_bl, inefficacy_bl,
    exhaustion_f01, cynicism_f01, inefficacy_f01)

# Paired t-tests for each dimension
t_exhaustion = t.test(overextended_ttest$exhaustion_bl, overextended_ttest$exhaustion_f01, paired = TRUE)
t_cynicism   = t.test(overextended_ttest$cynicism_bl,overextended_ttest$cynicism_f01,   paired = TRUE)
t_inefficacy = t.test(overextended_ttest$inefficacy_bl, overextended_ttest$inefficacy_f01, paired = TRUE)

bind_rows(
  tidy(t_exhaustion) |> mutate(dimension = "exhaustion"),
  tidy(t_cynicism)   |> mutate(dimension = "cynicism"),
  tidy(t_inefficacy) |> mutate(dimension = "inefficacy")) |>
  select(dimension, estimate, statistic, p.value, conf.low, conf.high) |> 
    kable(
    caption = "Overextended profile at baseline vs. follow-up 1",
    col.names = c("Dimension", "Mean Difference", "t", "p", "CI low", "CI high"),      digits = 3)

# Ineffective
ineffective_ttest = mbi_merged |>
  filter(bl_class == "Ineffective",
         f01_class == "Ineffective") |>
  select(case,exhaustion_bl, cynicism_bl, inefficacy_bl,
    exhaustion_f01, cynicism_f01, inefficacy_f01)

# Paired t-tests for each dimension
t_exhaustion = t.test(ineffective_ttest$exhaustion_bl, ineffective_ttest$exhaustion_f01, paired = TRUE)
t_cynicism   = t.test(ineffective_ttest$cynicism_bl,ineffective_ttest$cynicism_f01,   paired = TRUE)
t_inefficacy = t.test(ineffective_ttest$inefficacy_bl, ineffective_ttest$inefficacy_f01, paired = TRUE)

bind_rows(
  tidy(t_exhaustion) |> mutate(dimension = "exhaustion"),
  tidy(t_cynicism)   |> mutate(dimension = "cynicism"),
  tidy(t_inefficacy) |> mutate(dimension = "inefficacy")) |>
  select(dimension, estimate, statistic, p.value, conf.low, conf.high) |> 
    kable(
    caption = "Ineffective profile at baseline vs. follow-up 1",
    col.names = c("Dimension", "Mean Difference", "t", "p", "CI low", "CI high"),
    digits = 3)

# Engagement
engagement_ttest = mbi_merged |>
  filter(bl_class == "Engagement",
         f01_class == "Engagement") |>
  select(case,exhaustion_bl, cynicism_bl, inefficacy_bl,
    exhaustion_f01, cynicism_f01, inefficacy_f01)

# Paired t-tests for each dimension
t_exhaustion = t.test(engagement_ttest$exhaustion_bl, engagement_ttest$exhaustion_f01, paired = TRUE)
t_cynicism   = t.test(engagement_ttest$cynicism_bl,engagement_ttest$cynicism_f01,   paired = TRUE)
t_inefficacy = t.test(engagement_ttest$inefficacy_bl, engagement_ttest$inefficacy_f01, paired = TRUE)

bind_rows(
  tidy(t_exhaustion) |> mutate(dimension = "exhaustion"),
  tidy(t_cynicism)   |> mutate(dimension = "cynicism"),
  tidy(t_inefficacy) |> mutate(dimension = "inefficacy")) |>
  select(dimension, estimate, statistic, p.value, conf.low, conf.high) |> 
    kable(
    caption = "Engagement profile at baseline vs. follow-up 1",
    col.names = c("Dimension", "Mean Difference", "t", "p", "CI low", "CI high"),
    digits = 3)

```

**Pool moderate classes**

We will merge the three moderate classes - **Disengaged, Overextended,** and **Ineffective** - into one "burnout risk" group. Thereby, class sizes increase substantially, thereby enhancing predictability, robustness, and replicability of findings. We also hope that by pooling the three classes, more consistent patterns will emerge in terms of transitions between profiles.

```{r}
#| label: pool classes into one "burnout risk" group

mbi_merged3 = mbi_merged |> 
  mutate(bl_class = fct_recode(bl_class,"Burnout" = "Burnout","Risk" = "Disengaged","Risk" = "Ineffective", "Risk" = "Overextended", "Engagement" = "Engagement"))

mbi_merged3 = mbi_merged3 |> 
  mutate(f01_class = fct_recode(f01_class,"Burnout" = "Burnout","Risk" = "Disengaged","Risk" = "Ineffective", "Risk" = "Overextended", "Engagement" = "Engagement"))

mbi_merged3 = mbi_merged3 |> 
  mutate(f02_class = fct_recode(f02_class,"Burnout" = "Burnout","Risk" = "Disengaged","Risk" = "Ineffective", "Risk" = "Overextended", "Engagement" = "Engagement"))
```

## Mean plots per timepoint

```{r message=FALSE, warning=FALSE}
#| label: line plots bl
#| echo: false

individual_data = mbi_merged3 |>
  select(case, exhaustion_bl, cynicism_bl, inefficacy_bl, bl_class) |>
  pivot_longer(
    cols = c(exhaustion_bl, cynicism_bl, inefficacy_bl),
    names_to = "dimension",
    values_to = "score"
  ) |>
  mutate(dimension = factor(dimension, levels = c("exhaustion_bl", "cynicism_bl", "inefficacy_bl")))

line_plotsred3bl = mbi_merged3 |> 
  select(case, exhaustion_bl, cynicism_bl, inefficacy_bl, bl_class) |> 
  pivot_longer(
    cols = c(exhaustion_bl, inefficacy_bl, cynicism_bl),
    names_to = "dimension",
    values_to = "score") |> 
  group_by(bl_class, dimension) |> 
  summarise(mean_score = mean(score),
            .groups = "drop")

#specify colors
profile_colorsred3bl = c(
  "Burnout" = "midnightblue",
  "Risk" = "steelblue3",
  "Engagement" = "lightskyblue")

line_plotsred3bl = line_plotsred3bl |> 
  mutate(bl_class = factor(bl_class)) |> 
  mutate(dimension = factor(dimension,levels = 
                              c("exhaustion_bl", "cynicism_bl", "inefficacy_bl")))

fig1 = ggplot() +
  # Individual trajectories — plotted first so they appear behind mean lines
  geom_line(data = individual_data, aes(x = dimension, y = score, group = case),
            color = "gray25", linewidth = 0.4, alpha = 0.1) +
  # Mean lines on top
  geom_line(data = line_plotsred3bl, aes(x = dimension, y = mean_score, group = bl_class, 
                                          color = bl_class), linewidth = 1.2) +
  geom_point(data = line_plotsred3bl, aes(x = dimension, y = mean_score, group = bl_class, 
                                           color = bl_class), size = 3) +
  scale_color_manual(values = profile_colorsred3bl) +
  scale_x_discrete(labels = c(
    exhaustion_f02 = "Exhaustion",
    cynicism_f02   = "Cynicism",
    inefficacy_f02 = "Inefficacy")) +
  scale_y_continuous(limits = c(0, 6)) +
  labs(y = "Mean Score", x = "", color = "Profile") +
  theme_minimal(base_size = 9) 
plot(fig1)

#setwd("C:/Users/simon/Nextcloud2/Shared/LPA - Simon/04_longitudinal_analyses/plots")
#ggsave("bl_3_mean.png", fig1, width = 12, height = 10, bg = "white", units = "cm", dpi = 400)

```

```{r warning=FALSE, message=FALSE}
#| label: line plots f01
#| echo: false

individual_data = mbi_merged3 |>
  select(case, exhaustion_f01, cynicism_f01, inefficacy_f01, f01_class) |>
  pivot_longer(
    cols = c(exhaustion_f01, cynicism_f01, inefficacy_f01),
    names_to = "dimension",
    values_to = "score"
  ) |>
  mutate(dimension = factor(dimension, levels = c("exhaustion_f01", "cynicism_f01", "inefficacy_f01")))

line_plotsred3f01 = mbi_merged3 |> 
  select(case, exhaustion_f01, cynicism_f01, inefficacy_f01, f01_class) |> 
  pivot_longer(
    cols = c(exhaustion_f01, inefficacy_f01, cynicism_f01),
    names_to = "dimension",
    values_to = "score") |> 
  group_by(f01_class, dimension) |> 
  summarise(mean_score = mean(score),
            .groups = "drop")

#specify colors
profile_colorsred3f01 = c(
  "Burnout" = "midnightblue",
  "Risk" = "steelblue3",
  "Engagement" = "lightskyblue")

line_plotsred3f01 = line_plotsred3f01 |> 
  mutate(f02_class = factor(f01_class)) |> 
  mutate(dimension = factor(dimension,levels = 
                              c("exhaustion_f01", "cynicism_f01", "inefficacy_f01")))

fig1 = ggplot() +
  # Individual trajectories — plotted first so they appear behind mean lines
  geom_line(data = individual_data, aes(x = dimension, y = score, group = case),
            color = "gray25", linewidth = 0.4, alpha = 0.1) +
  # Mean lines on top
  geom_line(data = line_plotsred3f01, aes(x = dimension, y = mean_score, group = f01_class, 
                                          color = f01_class), linewidth = 1.2) +
  geom_point(data = line_plotsred3f01, aes(x = dimension, y = mean_score, group = f01_class, 
                                           color = f01_class), size = 3) +
  scale_color_manual(values = profile_colorsred3f01) +
  scale_x_discrete(labels = c(
    exhaustion_f02 = "Exhaustion",
    cynicism_f02   = "Cynicism",
    inefficacy_f02 = "Inefficacy")) +
  scale_y_continuous(limits = c(0, 6)) +
  labs(y = "Mean Score", x = "", color = "Profile") +
  theme_minimal(base_size = 9) 
plot(fig1)

#setwd("C:/Users/simon/Nextcloud2/Shared/LPA - Simon/04_longitudinal_analyses/plots")
#ggsave("f01_3_mean.png", fig1, width = 12, height = 10, bg = "white", units = "cm", dpi = 400)

```

```{r warning=FALSE, message=FALSE}
#| label: line plots f02
#| echo: false

individual_data = mbi_merged3 |>
  select(case, exhaustion_f02, cynicism_f02, inefficacy_f02, f02_class) |>
  pivot_longer(
    cols = c(exhaustion_f02, cynicism_f02, inefficacy_f02),
    names_to = "dimension",
    values_to = "score"
  ) |>
  mutate(dimension = factor(dimension, levels = c("exhaustion_f02", "cynicism_f02", "inefficacy_f02")))

line_plotsred3f02 = mbi_merged3 |> 
  select(case, exhaustion_f02, cynicism_f02, inefficacy_f02, f02_class) |> 
  pivot_longer(
    cols = c(exhaustion_f02, inefficacy_f02, cynicism_f02),
    names_to = "dimension",
    values_to = "score") |> 
  group_by(f02_class, dimension) |> 
  summarise(mean_score = mean(score),
            .groups = "drop")

#specify colors
profile_colorsred3f02 = c(
  "Burnout" = "midnightblue",
  "Risk" = "steelblue3",
  "Engagement" = "lightskyblue")

line_plotsred3f02 = line_plotsred3f02 |> 
  mutate(f02_class = factor(f02_class)) |> 
  mutate(dimension = factor(dimension,levels = 
                              c("exhaustion_f02", "cynicism_f02", "inefficacy_f02")))

fig1 = ggplot() +
  # Individual trajectories — plotted first so they appear behind mean lines
  geom_line(data = individual_data, aes(x = dimension, y = score, group = case),
            color = "gray25", linewidth = 0.4, alpha = 0.1) +
  # Mean lines on top
  geom_line(data = line_plotsred3f02, aes(x = dimension, y = mean_score, group = f02_class, 
                                          color = f02_class), linewidth = 1.2) +
  geom_point(data = line_plotsred3f02, aes(x = dimension, y = mean_score, group = f02_class, 
                                           color = f02_class), size = 3) +
  scale_color_manual(values = profile_colorsred3f02) +
  scale_x_discrete(labels = c(
    exhaustion_f02 = "Exhaustion",
    cynicism_f02   = "Cynicism",
    inefficacy_f02 = "Inefficacy")) +
  scale_y_continuous(limits = c(0, 6)) +
  labs(y = "Mean Score", x = "", color = "Profile") +
  theme_minimal(base_size = 9) 
plot(fig1)

#setwd("C:/Users/simon/Nextcloud2/Shared/LPA - Simon/04_longitudinal_analyses/plots")
#ggsave("f02_3_mean.png", fig1, width = 12, height = 10, bg = "white", units = "cm", dpi = 400)
```

## LTA

### Alluvial plot

```{r warning = FALSE, message = FALSE}
#| label: alluvial plot 3 classes
#| echo: false

all1 = ggplot(mbi_merged3, aes(y = 1, axis1 = bl_class, axis2 = f01_class, axis3 = f02_class)) +
  geom_alluvium(aes(fill = bl_class), width = 1/12) +
  geom_stratum(aes(fill = after_stat(stratum)), width = 1/12, alpha = 0.8,
    color = "grey30") +
  geom_label(stat = "stratum", aes(label = after_stat(stratum)), color = "black") +
  scale_x_discrete(limits = c("Baseline", "Follow-up 1", "Follow-up 2"), expand = c(.05, .05)) +
  scale_fill_manual(
    name = "Burnout profiles",values = c(
      "Burnout"      = "midnightblue",
      "Risk"   = "steelblue3",
      "Engagement" = "lightskyblue"),
    na.translate = FALSE) +
  ggtitle("Longitudinal Stability of Latent Burnout Profiles") +
  theme_minimal()

all1
#setwd("C:/Users/simon/Nextcloud2/Shared/LPA - Simon/04_longitudinal_analyses/plots")
#ggsave("alluvial_3.png", all1, width = 24, height = 16, bg = "white", units = "cm", dpi = 600)
```

### LTA-like approach

Quantify transitions with an LTA-like approach [@lund2025]

```{r}
#| label: LTA
#| echo: false

lta_data = mbi_merged3 |> 
  select(case, bl_class, f01_class, f02_class)

# Probability of being in "Burnout" at F01
fit_burnout = glm(f01_class == "Burnout" ~ bl_class - 1, 
                  family = binomial,
                  data = lta_data)

coefs_burnout = coef(fit_burnout)
probs_burnout = exp(coefs_burnout) / (1 + exp(coefs_burnout))

print("Probability of Burnout at F01 by BL class:")
print(round(probs_burnout, 3))

# Probability of being in "Risk" at session 2
fit_overextended = glm(f01_class == "Risk" ~ bl_class - 1, 
                       family = binomial, 
                       data = lta_data)
coefs_overextended = coef(fit_overextended)
probs_overextended = exp(coefs_overextended) / (1 + exp(coefs_overextended))

print("Probability of Risk at F01 by BL class:")
print(round(probs_overextended, 3))

# Probability of being in "Engagement" at F01

fit_engagement = glm(f01_class == "Engagement" ~ bl_class - 1, 
                     family = binomial, 
                     data = lta_data)
coefs_engagement = coef(fit_engagement)
probs_engagement = exp(coefs_engagement) / (1 + exp(coefs_engagement))

print("Probability of Engagement at F01 by BL class:")
print(round(probs_engagement, 3))
```

```{r}
#| echo: false
# Probability of being in "Burnout" at F02
fit_burnout = glm(f02_class == "Burnout" ~ f01_class - 1, 
                  family = binomial,
                  data = lta_data)

coefs_burnout = coef(fit_burnout)
probs_burnout = exp(coefs_burnout) / (1 + exp(coefs_burnout))

print("Probability of Burnout at F02 by BL class:")
print(round(probs_burnout, 3))

# Probability of being in "Risk" at session 3
fit_overextended = glm(f02_class == "Risk" ~ f01_class - 1, 
                       family = binomial, 
                       data = lta_data)
coefs_overextended = coef(fit_overextended)
probs_overextended = exp(coefs_overextended) / (1 + exp(coefs_overextended))

print("Probability of Risk at F02 by BL class:")
print(round(probs_overextended, 3))

# Probability of being in "Engagement" at F02

fit_engagement = glm(f02_class == "Engagement" ~ f01_class - 1, 
                     family = binomial, 
                     data = lta_data)
coefs_engagement = coef(fit_engagement)
probs_engagement = exp(coefs_engagement) / (1 + exp(coefs_engagement))

print("Probability of Engagement at F02 by BL class:")
print(round(probs_engagement, 3))
```

## ERI as predictor

```{r warning=FALSE, message=FALSE}
#| label: load respective data
#| echo: false

setwd("C:/Users/simon/Nextcloud2/Shared/LPA - Simon/04_longitudinal_analyses/data")
eri_bl = read.csv2("export_eri_bl_20240620.csv")
eri_follow1 = read.csv2("export_eri_fu01_20240620.csv")
eri_follow2 = read.csv2("export_eri_fu02_20240620.csv")

eri_bl = eri_bl |> 
  rename(case = ProbCode) |> 
  mutate(case = str_remove(case, "-bl$")) |> 
  group_by(case) |> 
  # reverse-code reward items  
  mutate(across(starts_with(c("ERI5", "ERI6", "ERI7")),
                ~ recode(as.character(.x),
                         "4"="1","3"="2","2"="3","1"="4"),
                .names="{.col}_2")) |> 
  mutate(across(ends_with("_2"), ~ suppressWarnings(as.numeric(.x))))

eri_bl = eri_bl |> 
  dplyr::select(case, ERI1_Eff, ERI2_Eff, ERI3_Eff, ERI4_Rew, ERI5_Rew_2, ERI6_Rew_2, ERI7_Rew_2, ERI8_Rew, ERI9_Rew, ERI10_Rew)
eri_bl = eri_bl |> 
  rename(ERI5_Rew = ERI5_Rew_2,
         ERI6_Rew = ERI6_Rew_2,
         ERI7_Rew = ERI7_Rew_2)

eri_bl = eri_bl |> 
  group_by(case) |> 
  mutate(
    effort = if (any(is.na(c_across(ends_with("Eff"))))) NA_real_
    else sum(c_across(ends_with("Eff"))),
    reward = if (any(is.na(c_across(ends_with("Rew"))))) NA_real_
    else sum(c_across(ends_with("Rew"))),
    err = (effort/reward) * (7/3))  |> 
  ungroup() 

eri_bl = eri_bl |> 
  dplyr::select(case, err) |> 
  rename(errbl = err)
#head(eri_bl)

eri_follow1 = eri_follow1 |> 
  rename(case = ProbCode) |> 
  mutate(case = str_remove(case, "-fu01$")) |> 
  group_by(case) |> 
  # reverse-code reward items  
  mutate(across(starts_with(c("ERI5", "ERI6", "ERI7")),
                ~ recode(as.character(.x),
                         "4"="1","3"="2","2"="3","1"="4"),
                .names="{.col}_2")) |> 
  mutate(across(ends_with("_2"), ~ suppressWarnings(as.numeric(.x))))

eri_follow1 = eri_follow1 |> 
  dplyr::select(case, ERI1_Eff, ERI2_Eff, ERI3_Eff, ERI4_Rew, ERI5_Rew_2, ERI6_Rew_2, ERI7_Rew_2, ERI8_Rew, ERI9_Rew, ERI10_Rew)
eri_follow1 = eri_follow1 |> 
  rename(ERI5_Rew = ERI5_Rew_2,
         ERI6_Rew = ERI6_Rew_2,
         ERI7_Rew = ERI7_Rew_2)

eri_follow1 = eri_follow1 |> 
  group_by(case) |> 
  mutate(
    effort = if (any(is.na(c_across(ends_with("Eff"))))) NA_real_
    else sum(c_across(ends_with("Eff"))),
    reward = if (any(is.na(c_across(ends_with("Rew"))))) NA_real_
    else sum(c_across(ends_with("Rew"))),
    err = (effort/reward) * (7/3))  |> 
  ungroup() 

eri_follow1 = eri_follow1 |> 
  dplyr::select(case, err) |> 
  rename(err1 = err)
#head(eri_follow1)

eri_follow2 = eri_follow2 |> 
  rename(case = ProbCode) |> 
  mutate(case = str_remove(case, "-fu02$")) |> 
  group_by(case) |> 
  # reverse-code reward items  
  mutate(across(starts_with(c("ERI5", "ERI6", "ERI7")),
                ~ recode(as.character(.x),
                         "4"="1","3"="2","2"="3","1"="4"),
                .names="{.col}_2")) |> 
  mutate(across(ends_with("_2"), ~ suppressWarnings(as.numeric(.x))))

eri_follow2 = eri_follow2 |> 
  dplyr::select(case, ERI1_Eff, ERI2_Eff, ERI3_Eff, ERI4_Rew, ERI5_Rew_2, ERI6_Rew_2, ERI7_Rew_2, ERI8_Rew, ERI9_Rew, ERI10_Rew) |> 
  rename(ERI5_Rew = ERI5_Rew_2,
         ERI6_Rew = ERI6_Rew_2,
         ERI7_Rew = ERI7_Rew_2)

eri_follow2 = eri_follow2 |> 
  group_by(case) |> 
  mutate(
    effort = if (any(is.na(c_across(ends_with("Eff"))))) NA_real_
    else sum(c_across(ends_with("Eff"))),
    reward = if (any(is.na(c_across(ends_with("Rew"))))) NA_real_
    else sum(c_across(ends_with("Rew"))),
    err = (effort/reward) * (7/3))  |> 
  ungroup() 

eri_follow2 = eri_follow2 |> 
  dplyr::select(case, err) |> 
  rename(err2 = err)
#head(eri_follow2)

```

**Descriptive differences in ERR between profiles** at different timepoints:

```{r}

lta_data_eri3 |> 
  group_by(bl_class) |> 
  summarise(
    mean_errbl = mean(errbl, na.rm = TRUE))

anovaresult = aov(errbl ~ bl_class, data = lta_data_eri3)
summary(anovaresult)
TukeyHSD(anovaresult)

lta_data_eri3 |> 
  group_by(f01_class) |> 
  summarise(
    mean_err1 = mean(err1, na.rm = TRUE))

anovaresult = aov(err1 ~ f01_class, data = lta_data_eri3)
summary(anovaresult)
TukeyHSD(anovaresult)

lta_data_eri3 |> 
  group_by(f02_class) |> 
  summarise(
    mean_err2 = mean(err2, na.rm = TRUE))

anovaresult = aov(err2 ~ f02_class, data = lta_data_eri3)
summary(anovaresult)
TukeyHSD(anovaresult)
```

Engagement differs significantly from Burnout and Risk groups in terms of ERR values. At Follow-up 1 and 2, Burnout and Risk groups' ERR values also differ significantly.

... of class membership at BL/F01/F02. We will use a multinomial logistic approach suggested by [@nagin2010] which closely resembles BCH analyses [@bakk2021a] but can be implemented more easily. By incorporating regression weights for the 5-class-solution, we account for the classification uncertainty inherent to LPA, expressed by posterior probabilities of assignment.

```{r}
#| label: ERI predicting class membership at various timepoints, 5 classes
#| echo: false

lta_data_eri5 = mbi_merged |> 
  select(case, bl_class, f01_class, f02_class, bl_pprob, f1_pprob, f2_pprob)

lta_data_eri5 = left_join(lta_data_eri5, eri_bl, by = "case")
lta_data_eri5 = left_join(lta_data_eri5, eri_follow1, by = "case")
lta_data_eri5 = left_join(lta_data_eri5, eri_follow2, by = "case")
head(lta_data_eri5)
```

### Class memberships, 5-profile solution

ERI predicting class memberships at **baseline.** We predict class membership with 1-unit increases in the Effort-Reward Ratio (which equals a massive deterioration). This ratio usually ranges from roughly **0.3 to 2.0**, meaning the full scale spans less than 2 units. A one-unit increase therefore represents a **very large, near-maximal shift** in work stress — essentially moving from the healthy to the pathological end of the scale. Large ORs for such a big increment are expected and theoretically coherent. We therefore also provide the OR values for 0.1-unit increases, which is much more digestible.

```{r}

lta_data_eri5bl = lta_data_eri5 |> 
  select(bl_class, bl_pprob, errbl)

#set reference level
lta_data_eri5bl$bl_class = relevel(as.factor(lta_data_eri5bl$bl_class), ref = "Engagement")

# Class as outcome, err as predictor, weights = posterior probabilities
mod_err = multinom(bl_class ~ errbl, data = lta_data_eri5bl, weights = bl_pprob, trace = FALSE)
summary(mod_err)
baseline = summary(mod_err)

# extract coefficients + SEs
coef_mat = baseline$coefficients
se_mat   = baseline$standard.errors

##################
#Burnout
# lower and upper CI on log-odds scale
lower_log_burnout = coef_mat[1,2] - 1.96 * se_mat[1,2]
upper_log_burnout = coef_mat[1,2] + 1.96 * se_mat[1,2]

# exponentiate: OR CI
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
ORerr_burnout = exp(coef_mat[1,2])

# Wald z-test
z_stats_burnout = coef_mat[1,2] / se_mat[1,2]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

####################
#Disengaged
# lower and upper CI on log-odds scale
lower_log_disengaged = coef_mat[2,2] - 1.96 * se_mat[2,2]
upper_log_disengaged = coef_mat[2,2] + 1.96 * se_mat[2,2]

# exponentiate: OR CI
lower_OR_disengaged = exp(lower_log_disengaged)
upper_OR_disengaged = exp(upper_log_disengaged)

# odds ratios
ORerr_disengaged = exp(coef_mat[2,2])

# Wald z-test
z_stats_disengaged = coef_mat[2,2] / se_mat[2,2]
p_values_disengaged = 2 * (1 - pnorm(abs(z_stats_disengaged)))

#####################
#Overextended
# lower and upper CI on log-odds scale
lower_log_overextended = coef_mat[4,2] - 1.96 * se_mat[4,2]
upper_log_overextended = coef_mat[4,2] + 1.96 * se_mat[4,2]

# exponentiate: OR CI
lower_OR_overextended = exp(lower_log_overextended)
upper_OR_overextended = exp(upper_log_overextended)

# odds ratios
ORerr_overextended = exp(coef_mat[4,2])

# Wald z-test
z_stats_overextended = coef_mat[4,2] / se_mat[4,2]
p_values_overextended = 2 * (1 - pnorm(abs(z_stats_overextended)))

#####################
#Ineffective

lower_log_ineffective = coef_mat[3,2] - 1.96 * se_mat[3,2]
upper_log_ineffective = coef_mat[3,2] + 1.96 * se_mat[3,2]

# exponentiate: OR CI
lower_OR_ineffective = exp(lower_log_ineffective)
upper_OR_ineffective = exp(upper_log_ineffective)

# odds ratios
ORerr_ineffective = exp(coef_mat[3,2])

# Wald z-test
z_stats_ineffective = coef_mat[3,2] / se_mat[3,2]
p_values_ineffective = 2 * (1 - pnorm(abs(z_stats_ineffective)))

```

**Higher ERR scores** by 1 unit (i.e., more efforts required per reward) at baseline were associated with 1500% higher odds (*31% higher odds for 0.1-unit-increases*) of being in **Burnout** compared to **Engagement** (OR = `{r} ORerr_burnout` , 95% CI\[`{r} lower_OR_burnout`,`{r} upper_OR_burnout`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_burnout`*,* *p* = `{r} p_values_burnout`.

**Higher ERR scores** by 1 unit (i.e., more efforts required per reward) at baseline were associated with 2000% higher odds (*35% higher odds for 0.1-unit-increases*) of being in **Disengaged** compared to **Engagement** (OR = `{r} ORerr_disengaged` , 95% CI\[`{r} lower_OR_disengaged`,`{r} upper_OR_disengaged`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_disengaged`*,* *p* = `{r} p_values_disengaged`.

**Higher ERR scores** by 1 unit (i.e., more efforts required per reward) at baseline were associated with 1100% higher odds (*27% higher odds for 0.1-unit-increases*) of being in **Overextended** compared to **Engagement** (OR = `{r} ORerr_overextended` , 95% CI\[`{r} lower_OR_overextended`,`{r} upper_OR_overextended`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_overextended`*,* *p* = `{r} p_values_overextended`.

**Higher ERR scores** by 1 unit (i.e., more efforts required per reward) at baseline were associated with 600% higher odds (*20% higher odds for 0.1-unit-increases*) of being in **Ineffective** compared to **Engagement** (OR = `{r} ORerr_ineffective` , 95% CI\[`{r} lower_OR_ineffective`,`{r} upper_OR_ineffective`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_ineffective`*,* *p* = `{r} p_values_ineffective`.

ERI predicting class memberships at **Follow-up 1:**

```{r}

lta_data_eri5f1 = lta_data_eri5 |> 
  select(f01_class, f1_pprob, err1)

#set reference level
lta_data_eri5f1$f01_class = relevel(as.factor(lta_data_eri5f1$f01_class), ref = "Engagement")

# Class as outcome, err as predictor, weights = posterior probabilities
mod_err = multinom(f01_class ~ err1, data = lta_data_eri5f1, weights = f1_pprob, trace = FALSE)
summary(mod_err)
baseline = summary(mod_err)

# extract coefficients + SEs
coef_mat = baseline$coefficients
se_mat   = baseline$standard.errors

##################
#Burnout
# lower and upper CI on log-odds scale
lower_log_burnout = coef_mat[1,2] - 1.96 * se_mat[1,2]
upper_log_burnout = coef_mat[1,2] + 1.96 * se_mat[1,2]

# exponentiate: OR CI
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
ORerr_burnout = exp(coef_mat[1,2])

# Wald z-test
z_stats_burnout = coef_mat[1,2] / se_mat[1,2]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

####################
#Disengaged
# lower and upper CI on log-odds scale
lower_log_disengaged = coef_mat[2,2] - 1.96 * se_mat[2,2]
upper_log_disengaged = coef_mat[2,2] + 1.96 * se_mat[2,2]

# exponentiate: OR CI
lower_OR_disengaged = exp(lower_log_disengaged)
upper_OR_disengaged = exp(upper_log_disengaged)

# odds ratios
ORerr_disengaged = exp(coef_mat[2,2])

# Wald z-test
z_stats_disengaged = coef_mat[2,2] / se_mat[2,2]
p_values_disengaged = 2 * (1 - pnorm(abs(z_stats_disengaged)))

#####################
#Overextended
# lower and upper CI on log-odds scale
lower_log_overextended = coef_mat[4,2] - 1.96 * se_mat[4,2]
upper_log_overextended = coef_mat[4,2] + 1.96 * se_mat[4,2]

# exponentiate: OR CI
lower_OR_overextended = exp(lower_log_overextended)
upper_OR_overextended = exp(upper_log_overextended)

# odds ratios
ORerr_overextended = exp(coef_mat[4,2])

# Wald z-test
z_stats_overextended = coef_mat[4,2] / se_mat[4,2]
p_values_overextended = 2 * (1 - pnorm(abs(z_stats_overextended)))

#####################
#Ineffective

lower_log_ineffective = coef_mat[3,2] - 1.96 * se_mat[3,2]
upper_log_ineffective = coef_mat[3,2] + 1.96 * se_mat[3,2]

# exponentiate: OR CI
lower_OR_ineffective = exp(lower_log_ineffective)
upper_OR_ineffective = exp(upper_log_ineffective)

# odds ratios
ORerr_ineffective = exp(coef_mat[3,2])

# Wald z-test
z_stats_ineffective = coef_mat[3,2] / se_mat[3,2]
p_values_ineffective = 2 * (1 - pnorm(abs(z_stats_ineffective)))
```

**Higher ERR scores** by 1 unit (i.e., more efforts required per reward) at follow-up 1 were associated with 4200% higher odds (*45% higher odds for 0.1-unit-increases*) of being in **Burnout** compared to **Engagement** (OR = `{r} ORerr_burnout` , 95% CI\[`{r} lower_OR_burnout`,`{r} upper_OR_burnout`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_burnout`*,* *p* = `{r} p_values_burnout`.

**Higher ERR scores** by 1 unit (i.e., more efforts required per reward) at follow-up 1 were associated with 3800% higher odds (*44% higher odds for 0.1-unit-increases*) of being in **Disengaged** compared to **Engagement** (OR = `{r} ORerr_disengaged` , 95% CI\[`{r} lower_OR_disengaged`,`{r} upper_OR_disengaged`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_disengaged`*,* *p* = `{r} p_values_disengaged`.

**Higher ERR scores** by 1 unit (i.e., more efforts required per reward) at follow-up 1 were associated with 1400% higher odds (*30% higher odds for 0.1-unit-increases*) of being in **Overextended** compared to **Engagement** (OR = `{r} ORerr_overextended` , 95% CI\[`{r} lower_OR_overextended`,`{r} upper_OR_overextended`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_overextended`*,* *p* = `{r} p_values_overextended`.

**Higher ERR scores** by 1 unit (i.e., more efforts required per reward) at follow-up 1 were associated with 500% higher odds (*19% higher odds for 0.1-unit-increases*) of being in **Ineffective** compared to **Engagement** (OR = `{r} ORerr_ineffective` , 95% CI\[`{r} lower_OR_ineffective`,`{r} upper_OR_ineffective`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_ineffective`*,* *p* = `{r} p_values_ineffective`.

### Class memberships, pooled (3-profile) solution

We cannot include posterior probabilities as regression weights for the pooled profile solution, as the pooling of three classes was a theory-driven, not data-driven, approach.

```{r}
#| label: ERI predicting class membership at various timepoints, pooled solution
#| echo: false

lta_data_eri3 = left_join(lta_data, eri_bl, by = "case")
lta_data_eri3 = left_join(lta_data_eri3, eri_follow1, by = "case")
lta_data_eri3 = left_join(lta_data_eri3, eri_follow2, by = "case")
head(lta_data_eri3)

```

ERI predicting class membership at **baseline, pooled burnout profiles**:

```{r}

lta_data_eri3bl = lta_data_eri3 |> 
  select(bl_class, errbl)

#set reference level
lta_data_eri3bl$bl_class = relevel(as.factor(lta_data_eri3bl$bl_class), ref = "Engagement")

# Class as outcome, err as predictor, weights = posterior probabilities
mod_err = multinom(bl_class ~ errbl, data = lta_data_eri3bl, trace = FALSE)
summary(mod_err)
baseline = summary(mod_err)

# extract coefficients + SEs
coef_mat = baseline$coefficients
se_mat   = baseline$standard.errors

##################
#Burnout
# lower and upper CI on log-odds scale
lower_log_burnout = coef_mat[1,2] - 1.96 * se_mat[1,2]
upper_log_burnout = coef_mat[1,2] + 1.96 * se_mat[1,2]

# exponentiate: OR CI
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
ORerr_burnout = exp(coef_mat[1,2])

# Wald z-test
z_stats_burnout = coef_mat[1,2] / se_mat[1,2]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

####################
#Risk
# lower and upper CI on log-odds scale
lower_log_risk = coef_mat[2,2] - 1.96 * se_mat[2,2]
upper_log_risk = coef_mat[2,2] + 1.96 * se_mat[2,2]

# exponentiate: OR CI
lower_OR_risk = exp(lower_log_risk)
upper_OR_risk = exp(upper_log_risk)

# odds ratios
ORerr_risk = exp(coef_mat[2,2])

# Wald z-test
z_stats_risk = coef_mat[2,2] / se_mat[2,2]
p_values_risk = 2 * (1 - pnorm(abs(z_stats_risk)))
```

**Higher ERR scores** by 1 unit (i.e., more efforts required per reward) at **baseline** were associated with 1300% higher odds (*30% higher odds for 0.1-unit-increases*) of being in **Burnout** compared to **Engagement** (OR = `{r} ORerr_burnout` , 95% CI\[`{r} lower_OR_burnout`,`{r} upper_OR_burnout`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_burnout`*,* *p* = `{r} p_values_burnout`.

**Higher ERR scores** by 1 unit (i.e., more efforts required per reward) at **baseline** were associated with 1000% higher odds (*26% higher odds for 0.1-unit-increases*) of being in **Risk** compared to **Engagement** (OR = `{r} ORerr_risk` , 95% CI\[`{r} lower_OR_risk`,`{r} upper_OR_risk`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_risk`*,* *p* = `{r} p_values_risk`.

ERI predicting class memberships at **follow-up 1, pooled burnout profiles**:

```{r}

lta_data_eri3f1 = lta_data_eri3 |> 
  select(f01_class, err1)

#set reference level
lta_data_eri3f1$f01_class = relevel(as.factor(lta_data_eri3f1$f01_class), ref = "Engagement")

# Class as outcome, err as predictor, weights = posterior probabilities
mod_err = multinom(f01_class ~ err1, data = lta_data_eri3f1, trace = FALSE)
summary(mod_err)
baseline = summary(mod_err)

# extract coefficients + SEs
coef_mat = baseline$coefficients
se_mat   = baseline$standard.errors

##################
#Burnout
# lower and upper CI on log-odds scale
lower_log_burnout = coef_mat[1,2] - 1.96 * se_mat[1,2]
upper_log_burnout = coef_mat[1,2] + 1.96 * se_mat[1,2]

# exponentiate: OR CI
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
ORerr_burnout = exp(coef_mat[1,2])

# Wald z-test
z_stats_burnout = coef_mat[1,2] / se_mat[1,2]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

####################
#Risk
# lower and upper CI on log-odds scale
lower_log_risk = coef_mat[2,2] - 1.96 * se_mat[2,2]
upper_log_risk = coef_mat[2,2] + 1.96 * se_mat[2,2]

# exponentiate: OR CI
lower_OR_risk = exp(lower_log_risk)
upper_OR_risk = exp(upper_log_risk)

# odds ratios
ORerr_risk = exp(coef_mat[2,2])

# Wald z-test
z_stats_risk = coef_mat[2,2] / se_mat[2,2]
p_values_risk = 2 * (1 - pnorm(abs(z_stats_risk)))
```

**Higher ERR scores** by 1 unit (i.e., more efforts required per reward) at **follow-up 1** were associated with 2500% higher odds (*38% higher odds for 0.1-unit-increases*) of being in **Burnout** compared to **Engagement** (OR = `{r} ORerr_burnout` , 95% CI\[`{r} lower_OR_burnout`,`{r} upper_OR_burnout`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_burnout`*,* *p* = `{r} p_values_burnout`.

**Higher ERR scores** by 1 unit (i.e., more efforts required per reward) at **follow-up 1** were associated with 1250% higher odds (*29% higher odds for 0.1-unit-increases*) of being in **Risk** compared to **Engagement** (OR = `{r} ORerr_risk` , 95% CI\[`{r} lower_OR_risk`,`{r} upper_OR_risk`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_risk`*,* *p* = `{r} p_values_risk`.

### Transitions, pooled (3-class) solution

ERI predicting the transitions between **pooled burnout profiles Baseline -\> Follow-up 1**

```{r}
#| label: ERI predicting BL->F01 pooled

# Transition 1: Burnout0 → Burnout1 
fit_to_Burnout1 = glm(
  f01_class == "Burnout" ~ bl_class + bl_class:err1 - err1,
  family = binomial,
  data = lta_data_eri3)

summary(fit_to_Burnout1)

burnoutcoef = coef(summary(fit_to_Burnout1))
exp(burnoutcoef[4,1]) #1.7 (~70% increase in likelihood, ~5% with 0.1 units)

CI_low =  exp(burnoutcoef[4,1] - 1.96 * burnoutcoef[4,2]) #0.5
CI_high = exp(burnoutcoef[4,1] + 1.96 * burnoutcoef[4,2]) #5.9
p_val = burnoutcoef[4,4] #0.40

# Transition 2: Risk0 → Burnout1 
exp(burnoutcoef[5,1]) #2.6 (~160% increase in likelihood, ~10% with 0.1 units)

CI_low = exp(burnoutcoef[5,1] - 1.96 * burnoutcoef[5,2]) #1.3
CI_high = exp(burnoutcoef[5,1] + 1.96 * burnoutcoef[5,2]) #5.2
p_val = burnoutcoef[5,4] #0.01

# Transition 3: Engagement0 → Burnout1 
exp(burnoutcoef[6,1]) #18 (~1800% increase in likelihood, ~34% with 0.1 units)

CI_low = exp(burnoutcoef[6,1] - 1.96 * burnoutcoef[6,2]) #0.2
CI_high = exp(burnoutcoef[6,1] + 1.96 * burnoutcoef[6,2]) #1616
p_val = burnoutcoef[6,4] #0.21

# Transition 4: Burnout0 → Risk1 
fit_to_risk = glm(
  f01_class == "Risk" ~ bl_class + bl_class:err1 - err1,
  family = binomial,
  data = lta_data_eri3)

summary(fit_to_risk)

riskcoef = coef(summary(fit_to_risk))
exp(riskcoef[4,1]) #1.5 (~50% increase in likelihood, ~4% with 0.1 units)

CI_low = exp(riskcoef[4,1] - 1.96 * riskcoef[4,2]) #0.4
CI_high = exp(riskcoef[4,1] + 1.96 * riskcoef[4,2]) #5.0
p_val = riskcoef[4,4] #0.52

# Transition 5: Risk0 → Risk1 
exp(riskcoef[5,1]) #2.5 (~150% increase in likelihood, ~10% with 0.1 units)

CI_low = exp(riskcoef[5,1] - 1.96 * riskcoef[5,2]) #1.4
CI_high = exp(riskcoef[5,1] + 1.96 * riskcoef[5,2]) #4.5
p_val = riskcoef[5,4] #0.003

# Transition 6: Engagement0 → Risk1 
exp(riskcoef[6,1]) #7.8 (~780% increase in likelihood, ~23% with 0.1 units)

CI_low = exp(riskcoef[6,1] - 1.96 * riskcoef[6,2]) #3.9
CI_high = exp(riskcoef[6,1] + 1.96 * riskcoef[6,2]) #15.9
p_val = riskcoef[6,4] #0.00

# Transition 7: Burnout0 → Engagement1
fit_to_engagement = glm(
  f01_class == "Engagement" ~ bl_class + bl_class:err1 - err1,
  family = binomial,
  data = lta_data_eri3)

summary(fit_to_engagement)
engagementcoef = coef(summary(fit_to_engagement))
exp(engagementcoef[4,1]) #0.01 (~100% decrease in likelihood, ~40% with 0.1 units)

CI_low = exp(engagementcoef[4,1] - 1.96 * engagementcoef[4,2]) #0.00
CI_high = exp(engagementcoef[4,1] + 1.96 * engagementcoef[4,2]) #0.5
p_val = engagementcoef[4,4] #0.03

# Transition 8: Risk0 → Engagement1 
exp(engagementcoef[5,1]) #0.08 (~92% decrease in likelihood, ~22% with 0.1 units)

CI_low = exp(engagementcoef[5,1] - 1.96 * engagementcoef[5,2]) #0.0
CI_high = exp(engagementcoef[5,1] + 1.96 * engagementcoef[5,2]) #0.2
p_val = engagementcoef[5,4] #0.00

# Transition 9: Engagement0 → Engagement1 
exp(engagementcoef[6,1]) #0.12 (~88% decrease in likelihood, ~19% with 0.1 units)

CI_low = exp(engagementcoef[6,1] - 1.96 * engagementcoef[6,2]) #0.06
CI_high = exp(engagementcoef[6,1] + 1.96 * engagementcoef[6,2]) #0.25
p_val = engagementcoef[6,4] #0.00

```

... of transitions between **pooled burnout profiles F01 –\> F02:**

```{r}
#| label: ERI predicting F01->F02 pooled

# Transition 1: Burnout1 → Burnout2 
fit_to_Burnout1 = glm(
  f02_class == "Burnout" ~ f01_class + f01_class:err2 - err2,
  family = binomial,
  data = lta_data_eri3)

summary(fit_to_Burnout1)

burnoutcoef = coef(summary(fit_to_Burnout1))
#print(burnoutcoef[6,])
exp(burnoutcoef[4,1]) #2.9 (1.11exp(burnoutcoef[4,1]))

CI_low = exp(burnoutcoef[4,1] - 1.96 * burnoutcoef[4,2]) #0.65
CI_high = exp(burnoutcoef[4,1] + 1.96 * burnoutcoef[4,2]) #13
p_val = burnoutcoef[4,4] #0.16

# Transition 2: Risk1 → Burnout2 
burnoutcoef = coef(summary(fit_to_Burnout1))
#print(burnoutcoef[7,])
exp(burnoutcoef[5,1]) #5.1 (1.18 10%)

CI_low = exp(burnoutcoef[5,1] - 1.96 * burnoutcoef[5,2]) #3.1
CI_high = exp(burnoutcoef[5,1] + 1.96 * burnoutcoef[5,2]) #8.3
p_val = burnoutcoef[5,4] #0.00

# Transition 3: Engagement1 → Burnout2 
burnoutcoef = coef(summary(fit_to_Burnout1))
#print(burnoutcoef[9,])
exp(burnoutcoef[6,1]) #32.4 (1.41 10%)

CI_low = exp(burnoutcoef[6,1] - 1.96 * burnoutcoef[6,2]) #2.2
CI_high = exp(burnoutcoef[6,1] + 1.96 * burnoutcoef[6,2]) #471
p_val = burnoutcoef[6,4] #0.01

# Transition 4: Burnout1 → Risk2 
fit_to_Disengaged1 = glm(
  f02_class == "Risk" ~ f01_class + f01_class:err2 - err2,
  family = binomial,
  data = lta_data_eri3)

summary(fit_to_Disengaged1)

disengagedcoef = coef(summary(fit_to_Disengaged1))
#print(disengagedcoef[6,])
exp(disengagedcoef[4,1]) #0.66 (0.96 10%)

CI_low = exp(disengagedcoef[4,1] - 1.96 * disengagedcoef[4,2]) #0.1
CI_high = exp(disengagedcoef[4,1] + 1.96 * disengagedcoef[4,2]) #3.1
p_val = disengagedcoef[4,4] #0.60

# Transition 5: Risk1 → Risk2 
disengagedcoef = coef(summary(fit_to_Disengaged1))
#print(disengagedcoef[7,])
exp(disengagedcoef[5,1]) #0.67 (0.96 10%)

CI_low = exp(disengagedcoef[5,1] - 1.96 * disengagedcoef[5,2]) #0.5
CI_high = exp(disengagedcoef[5,1] + 1.96 * disengagedcoef[5,2]) #0.98
p_val = disengagedcoef[5,4] #0.04

# Transition 6: Engagement1 → Risk2 
disengagedcoef = coef(summary(fit_to_Disengaged1))
#print(disengagedcoef[9,])
exp(disengagedcoef[6,1]) #5.1 (1.18 10%)

CI_low = exp(disengagedcoef[6,1] - 1.96 * disengagedcoef[6,2]) #2.2
CI_high = exp(disengagedcoef[6,1] + 1.96 * disengagedcoef[6,2]) #11.9
p_val = disengagedcoef[6,4] #0.00

# Transition 7: Burnout1 → Engagement2 
fit_to_engagement = glm(
  f02_class == "Engagement" ~ f01_class + f01_class:err2 - err2,
  family = binomial,
  data = lta_data_eri3)

summary(fit_to_engagement)
engagementcoef = coef(summary(fit_to_engagement))
exp(engagementcoef[4,1]) #0.13 (0.81 10%)

CI_low = exp(engagementcoef[4,1] - 1.96 * engagementcoef[4,2]) #0.00
CI_high = exp(engagementcoef[4,1] + 1.96 * engagementcoef[4,2]) #2.2
p_val = engagementcoef[4,4] #0.16

# Transition 8: Risk1 → Engagement2 
exp(engagementcoef[5,1]) #0.12 (0.80 10%)

CI_low = exp(engagementcoef[5,1] - 1.96 * engagementcoef[5,2]) #0.1
CI_high = exp(engagementcoef[5,1] + 1.96 * engagementcoef[5,2]) #0.3
p_val = engagementcoef[5,4] #0.00

# Transition 9: Engagement1 → Engagement2 
exp(engagementcoef[6,1]) #0.15 (0.83 10%)

CI_low = exp(engagementcoef[6,1] - 1.96 * engagementcoef[6,2]) #0.06
CI_high = exp(engagementcoef[6,1] + 1.96 * engagementcoef[6,2]) #0.36
p_val = engagementcoef[6,4] #0.00
```

Does a difference variable containing (ERR at Follow-up 1 **minus** ERR at Baseline) predict transitions from baseline to follow-up 1?

```{r}

#| label: ERI difference variable predicting BL->F01 pooled

lta_data_eri3 = lta_data_eri3 |> 
  mutate(diff_f1minusbl = err1-errbl) 

# Transition 1: Burnout0 → Burnout1 
fit_to_Burnout1 = glm(
  f01_class == "Burnout" ~ bl_class + bl_class:diff_f1minusbl - diff_f1minusbl,
  family = binomial,
  data = lta_data_eri3)

summary(fit_to_Burnout1)

burnoutcoef = coef(summary(fit_to_Burnout1))
exp(burnoutcoef[4,1]) #4.1 (~400% increase in likelihood, ~15% with 0.1 units)

CI_low =  exp(burnoutcoef[4,1] - 1.96 * burnoutcoef[4,2]) #1.1
CI_high = exp(burnoutcoef[4,1] + 1.96 * burnoutcoef[4,2]) #16
p_val = burnoutcoef[4,4] #0.04

# Transition 2: Risk0 → Burnout1 
exp(burnoutcoef[5,1]) #1.04 (~4% increase in likelihood, ~0% with 0.1 units)

CI_low = exp(burnoutcoef[5,1] - 1.96 * burnoutcoef[5,2]) #0.4
CI_high = exp(burnoutcoef[5,1] + 1.96 * burnoutcoef[5,2]) #2.7
p_val = burnoutcoef[5,4] #0.9

# Transition 3: Engagement0 → Burnout1 
exp(burnoutcoef[6,1]) #1216 (~inf% increase in likelihood, ~200% with 0.1 units)

CI_low = exp(burnoutcoef[6,1] - 1.96 * burnoutcoef[6,2]) #0.8
CI_high = exp(burnoutcoef[6,1] + 1.96 * burnoutcoef[6,2]) #1887345
p_val = burnoutcoef[6,4] #0.06

# Transition 4: Burnout0 → Risk1 
fit_to_risk = glm(
  f01_class == "Risk" ~ bl_class + bl_class:diff_f1minusbl - diff_f1minusbl,
  family = binomial,
  data = lta_data_eri3)

summary(fit_to_risk)

riskcoef = coef(summary(fit_to_risk))
exp(riskcoef[4,1]) #0.69 (~31% decrease in likelihood, ~4% with 0.1 units)

CI_low = exp(riskcoef[4,1] - 1.96 * riskcoef[4,2]) #0.25
CI_high = exp(riskcoef[4,1] + 1.96 * riskcoef[4,2]) #1.9
p_val = riskcoef[4,4] #0.49

# Transition 5: Risk0 → Risk1 
exp(riskcoef[5,1]) #4.0 (~400% increase in likelihood, ~15% with 0.1 units)

CI_low = exp(riskcoef[5,1] - 1.96 * riskcoef[5,2]) #2.3
CI_high = exp(riskcoef[5,1] + 1.96 * riskcoef[5,2]) #7.2
p_val = riskcoef[5,4] #0.00

# Transition 6: Engagement0 → Risk1 
exp(riskcoef[6,1]) #2.61 (~260% increase in likelihood, ~10% with 0.1 units)

CI_low = exp(riskcoef[6,1] - 1.96 * riskcoef[6,2]) #1.2
CI_high = exp(riskcoef[6,1] + 1.96 * riskcoef[6,2]) #5.6
p_val = riskcoef[6,4] #0.01

# Transition 7: Burnout0 → Engagement1
fit_to_engagement = glm(
  f01_class == "Engagement" ~ bl_class + bl_class:diff_f1minusbl - diff_f1minusbl,
  family = binomial,
  data = lta_data_eri3)

summary(fit_to_engagement)
engagementcoef = coef(summary(fit_to_engagement))
exp(engagementcoef[4,1]) #0.13 (~87% decrease in likelihood, ~18% with 0.1 units)

CI_low = exp(engagementcoef[4,1] - 1.96 * engagementcoef[4,2]) #0.02
CI_high = exp(engagementcoef[4,1] + 1.96 * engagementcoef[4,2]) #0.82
p_val = engagementcoef[4,4] #0.03

# Transition 8: Risk0 → Engagement1 
exp(engagementcoef[5,1]) #0.17 (~83% decrease in likelihood, ~16% with 0.1 units)

CI_low = exp(engagementcoef[5,1] - 1.96 * engagementcoef[5,2]) #0.1
CI_high = exp(engagementcoef[5,1] + 1.96 * engagementcoef[5,2]) #0.3
p_val = engagementcoef[5,4] #0.00

# Transition 9: Engagement0 → Engagement1 
exp(engagementcoef[6,1]) #0.34 (~66% decrease in likelihood, ~10% with 0.1 units)

CI_low = exp(engagementcoef[6,1] - 1.96 * engagementcoef[6,2]) #0.16
CI_high = exp(engagementcoef[6,1] + 1.96 * engagementcoef[6,2]) #0.75
p_val = engagementcoef[6,4] #0.01

```

Does a difference variable (ERRfollowup2 **minus** ERRfollowup1) predict transitions from Follow-up 1 to Follow-up 2?

```{r}
#| label: ERI difference variable predicting F01->F02 pooled

lta_data_eri3 = lta_data_eri3 |> 
  mutate(diff_f2minusf1 = err2-err1) 

# Transition 1: Burnout0 → Burnout1 
fit_to_Burnout1 = glm(
  f02_class == "Burnout" ~ f01_class + f01_class:diff_f2minusf1 - diff_f2minusf1,
  family = binomial,
  data = lta_data_eri3)

summary(fit_to_Burnout1)

burnoutcoef = coef(summary(fit_to_Burnout1))
exp(burnoutcoef[4,1]) #1.8 (~80% increase in likelihood, ~6% with 0.1 units)

CI_low =  exp(burnoutcoef[4,1] - 1.96 * burnoutcoef[4,2]) #0.3
CI_high = exp(burnoutcoef[4,1] + 1.96 * burnoutcoef[4,2]) #12.8
p_val = burnoutcoef[4,4] #0.56

# Transition 2: Risk0 → Burnout1 
exp(burnoutcoef[5,1]) #2.5 (~250% increase in likelihood, ~10% with 0.1 units)

CI_low = exp(burnoutcoef[5,1] - 1.96 * burnoutcoef[5,2]) #1.4
CI_high = exp(burnoutcoef[5,1] + 1.96 * burnoutcoef[5,2]) #4.3
p_val = burnoutcoef[5,4] #0.00

# Transition 3: Engagement0 → Burnout1 
exp(burnoutcoef[6,1]) #33.8 (~3300% increase in likelihood, ~42% with 0.1 units)

CI_low = exp(burnoutcoef[6,1] - 1.96 * burnoutcoef[6,2]) #1.2
CI_high = exp(burnoutcoef[6,1] + 1.96 * burnoutcoef[6,2]) #965
p_val = burnoutcoef[6,4] #0.04

# Transition 4: Burnout0 → Risk1 
fit_to_risk = glm(
  f02_class == "Risk" ~ f01_class + f01_class:diff_f2minusf1 - diff_f2minusf1,
  family = binomial,
  data = lta_data_eri3)

summary(fit_to_risk)

riskcoef = coef(summary(fit_to_risk))
exp(riskcoef[4,1]) #0.61 (~40% decrease in likelihood, ~5% with 0.1 units)

CI_low = exp(riskcoef[4,1] - 1.96 * riskcoef[4,2]) #0.06
CI_high = exp(riskcoef[4,1] + 1.96 * riskcoef[4,2]) #5.6
p_val = riskcoef[4,4] #0.66

# Transition 5: Risk0 → Risk1 
exp(riskcoef[5,1]) #0.93 (~7% decrease in likelihood, ~1% with 0.1 units)

CI_low = exp(riskcoef[5,1] - 1.96 * riskcoef[5,2]) #0.6
CI_high = exp(riskcoef[5,1] + 1.96 * riskcoef[5,2]) #1.5
p_val = riskcoef[5,4] #0.75

# Transition 6: Engagement0 → Risk1 
exp(riskcoef[6,1]) #4.1 (~400% increase in likelihood, ~15% with 0.1 units)

CI_low = exp(riskcoef[6,1] - 1.96 * riskcoef[6,2]) #1.4
CI_high = exp(riskcoef[6,1] + 1.96 * riskcoef[6,2]) #12
p_val = riskcoef[6,4] #0.01

# Transition 7: Burnout0 → Engagement1
fit_to_engagement = glm(
  f02_class == "Engagement" ~ f01_class + f01_class:diff_f2minusf1 - diff_f2minusf1,
  family = binomial,
  data = lta_data_eri3)

summary(fit_to_engagement)
engagementcoef = coef(summary(fit_to_engagement))
exp(engagementcoef[4,1]) #0.63 (~37% decrease in likelihood, ~4% with 0.1 units)

CI_low = exp(engagementcoef[4,1] - 1.96 * engagementcoef[4,2]) #0.03
CI_high = exp(engagementcoef[4,1] + 1.96 * engagementcoef[4,2]) #12
p_val = engagementcoef[4,4] #0.76

# Transition 8: Risk0 → Engagement1 
exp(engagementcoef[5,1]) #0.34 (~66% decrease in likelihood, ~10% with 0.1 units)

CI_low = exp(engagementcoef[5,1] - 1.96 * engagementcoef[5,2]) #0.18
CI_high = exp(engagementcoef[5,1] + 1.96 * engagementcoef[5,2]) #0.66
p_val = engagementcoef[5,4] #0.00

# Transition 9: Engagement0 → Engagement1 
exp(engagementcoef[6,1]) #0.20 (~80% decrease in likelihood, ~15% with 0.1 units)

CI_low = exp(engagementcoef[6,1] - 1.96 * engagementcoef[6,2]) #0.07
CI_high = exp(engagementcoef[6,1] + 1.96 * engagementcoef[6,2]) #0.58
p_val = engagementcoef[6,4] #0.00
```

## Burnout as predictor

```{r warning = FALSE, message = FALSE}
#| echo: false

setwd("C:/Users/simon/Nextcloud2/Shared/LPA - Simon/04_longitudinal_analyses/data")
burnout_bl = read_xlsx("Baseline - Burnout.xlsx")
burnout_follow1 = read_xlsx("FU01 - Burnout.xlsx")
burnout_follow2 = read_xlsx("FU02 - Burnout.xlsx")

burnout_bl = burnout_bl |> 
  rename(case = ProbCode) |> 
  mutate(case = str_remove(case, "-bl$")) 

burnout_bl = burnout_bl |> 
  dplyr::select(case, burnout_diagnosed) |> 
  mutate(burnout_diagnosed = factor(burnout_diagnosed)) |> 
  rename(burnout_bl = burnout_diagnosed)

burnout_follow1 = burnout_follow1 |> 
  rename(case = ProbCode) |> 
  mutate(case = str_remove(case, "-fu01$")) 

burnout_follow1 = burnout_follow1 |> 
  dplyr::select(case, burnout_diagnosed) |> 
  mutate(burnout_diagnosed = factor(burnout_diagnosed)) |> 
  rename(burnout_f01 = burnout_diagnosed)

burnout_follow2 = burnout_follow2 |> 
  rename(case = ProbCode) |> 
  mutate(case = str_remove(case, "-fu02$")) 

burnout_follow2 = burnout_follow2 |> 
  dplyr::select(case, burnout_diagnosed) |> 
  mutate(burnout_diagnosed = factor(burnout_diagnosed)) |> 
  rename(burnout_f02 = burnout_diagnosed)

burnout_data = left_join(burnout_bl, burnout_follow1, by = "case")
burnout_data = left_join(burnout_data, burnout_follow2, by = "case")

lta_data_burnout = left_join(lta_data, burnout_data, by = "case")
```

**Descriptive differences in Burnout frequency** between profiles:

```{r}
table_data_bl = table(lta_data_burnout$bl_class, lta_data_burnout$burnout_bl)
table_data_bl
chisq.test(table_data_bl)
```

```{r}
table_data_f01 = table(lta_data_burnout$f01_class, lta_data_burnout$burnout_f01)
table_data_f01
chisq.test(table_data_f01)
```

`{r} (table_data_f01[1,2]/(table_data_f01[1,1] + table_data_f01[1,2]))*100` % of participants in the Burnout profile have been diagnosed with Burnout at Follow-up 1 (as opposed to `{r} (table_data_bl[1,2]/(table_data_bl[1,1]) + table_data_bl[1,2])*100` at Baseline!)

```{r}
table_data_f02 = table(lta_data_burnout$f02_class, lta_data_burnout$burnout_f02)
table_data_f02
chisq.test(table_data_f02)
```

### Class memberships, pooled (3-profile) solution

```{r}
#| label: Burnout predicting class memberships

lta_data_burnoutbl = lta_data_burnout |> 
  select(bl_class, burnout_bl)

#set reference level
lta_data_burnoutbl$bl_class = relevel(as.factor(lta_data_burnoutbl$bl_class), ref = "Engagement")
lta_data_burnoutbl$burnout_bl = relevel(as.factor(lta_data_burnoutbl$burnout_bl), ref = "0")

# Class as outcome, err as predictor, weights = posterior probabilities
mod_err = multinom(bl_class ~ burnout_bl, data = lta_data_burnoutbl, trace = FALSE)
summary(mod_err)
baseline = summary(mod_err)

# extract coefficients + SEs
coef_mat = baseline$coefficients
se_mat   = baseline$standard.errors

##################
#Burnout
# lower and upper CI on log-odds scale
lower_log_burnout = coef_mat[1,2] - 1.96 * se_mat[1,2]
upper_log_burnout = coef_mat[1,2] + 1.96 * se_mat[1,2]

# exponentiate: OR CI
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
ORerr_burnout = exp(coef_mat[1,2])

# Wald z-test
z_stats_burnout = coef_mat[1,2] / se_mat[1,2]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

####################
#Risk
# lower and upper CI on log-odds scale
lower_log_risk = coef_mat[2,2] - 1.96 * se_mat[2,2]
upper_log_risk = coef_mat[2,2] + 1.96 * se_mat[2,2]

# exponentiate: OR CI
lower_OR_risk = exp(lower_log_risk)
upper_OR_risk = exp(upper_log_risk)

# odds ratios
ORerr_risk = exp(coef_mat[2,2])

# Wald z-test
z_stats_risk = coef_mat[2,2] / se_mat[2,2]
p_values_risk = 2 * (1 - pnorm(abs(z_stats_risk)))
```

**Having been diagnosed with Burnout** at **baseline** was associated with 140% higher odds of being in **Burnout** compared to **Engagement** (OR = `{r} ORerr_burnout` , 95% CI\[`{r} lower_OR_burnout`,`{r} upper_OR_burnout`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_burnout`*,* *p* = `{r} p_values_burnout`.

**Having been diagnosed with Burnout** at **baseline** was associated with 160% higher odds of being in **Risk** compared to **Engagement** (OR = `{r} ORerr_risk` , 95% CI\[`{r} lower_OR_risk`,`{r} upper_OR_risk`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_risk`*,* *p* = `{r} p_values_risk`.

ERI predicting class memberships at **follow-up 1, pooled burnout profiles**:

```{r}
lta_data_burnoutf1 = lta_data_burnout |> 
  select(f01_class, burnout_f01)

#set reference level
lta_data_burnoutf1$f01_class = relevel(as.factor(lta_data_burnoutf1$f01_class), ref = "Engagement")
lta_data_burnoutf1$burnout_f01 = relevel(as.factor(lta_data_burnoutf1$burnout_f01), ref = "0")

# Class as outcome, err as predictor, weights = posterior probabilities
mod_err = multinom(f01_class ~ burnout_f01, data = lta_data_burnoutf1, trace = FALSE)
summary(mod_err)
baseline = summary(mod_err)

# extract coefficients + SEs
coef_mat = baseline$coefficients
se_mat   = baseline$standard.errors

##################
#Burnout
# lower and upper CI on log-odds scale
lower_log_burnout = coef_mat[1,2] - 1.96 * se_mat[1,2]
upper_log_burnout = coef_mat[1,2] + 1.96 * se_mat[1,2]

# exponentiate: OR CI
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
ORerr_burnout = exp(coef_mat[1,2])

# Wald z-test
z_stats_burnout = coef_mat[1,2] / se_mat[1,2]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

####################
#Risk
# lower and upper CI on log-odds scale
lower_log_risk = coef_mat[2,2] - 1.96 * se_mat[2,2]
upper_log_risk = coef_mat[2,2] + 1.96 * se_mat[2,2]

# exponentiate: OR CI
lower_OR_risk = exp(lower_log_risk)
upper_OR_risk = exp(upper_log_risk)

# odds ratios
ORerr_risk = exp(coef_mat[2,2])

# Wald z-test
z_stats_risk = coef_mat[2,2] / se_mat[2,2]
p_values_risk = 2 * (1 - pnorm(abs(z_stats_risk)))
```

**Having been diagnosed with Burnout** at **follow-up 1** was associated with 550% higher odds of being in **Burnout** compared to **Engagement** (OR = `{r} ORerr_burnout` , 95% CI\[`{r} lower_OR_burnout`,`{r} upper_OR_burnout`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_burnout`*,* *p* = `{r} p_values_burnout`.

**Having been diagnosed with Burnout** at **follow-up 1** was associated with 180% higher odds of being in **Risk** compared to **Engagement** (OR = `{r} ORerr_risk` , 95% CI\[`{r} lower_OR_risk`,`{r} upper_OR_risk`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_risk`*,* *p* = `{r} p_values_risk`.

ERI predicting class memberships at **follow-up 2, pooled burnout profiles**:

```{r}
lta_data_burnoutf2 = lta_data_burnout |> 
  select(f02_class, burnout_f02)

#set reference level
lta_data_burnoutf2$f02_class = relevel(as.factor(lta_data_burnoutf2$f02_class), ref = "Engagement")
lta_data_burnoutf2$burnout_f02 = relevel(as.factor(lta_data_burnoutf2$burnout_f02), ref = "0")

# Class as outcome, err as predictor, weights = posterior probabilities
mod_err = multinom(f02_class ~ burnout_f02, data = lta_data_burnoutf2, trace = FALSE)
summary(mod_err)
baseline = summary(mod_err)

# extract coefficients + SEs
coef_mat = baseline$coefficients
se_mat   = baseline$standard.errors

##################
#Burnout
# lower and upper CI on log-odds scale
lower_log_burnout = coef_mat[1,2] - 1.96 * se_mat[1,2]
upper_log_burnout = coef_mat[1,2] + 1.96 * se_mat[1,2]

# exponentiate: OR CI
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
ORerr_burnout = exp(coef_mat[1,2])

# Wald z-test
z_stats_burnout = coef_mat[1,2] / se_mat[1,2]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

####################
#Risk
# lower and upper CI on log-odds scale
lower_log_risk = coef_mat[2,2] - 1.96 * se_mat[2,2]
upper_log_risk = coef_mat[2,2] + 1.96 * se_mat[2,2]

# exponentiate: OR CI
lower_OR_risk = exp(lower_log_risk)
upper_OR_risk = exp(upper_log_risk)

# odds ratios
ORerr_risk = exp(coef_mat[2,2])

# Wald z-test
z_stats_risk = coef_mat[2,2] / se_mat[2,2]
p_values_risk = 2 * (1 - pnorm(abs(z_stats_risk)))
```

**Having been diagnosed with Burnout** until **follow-up 2** was associated with 340% higher odds of being in **Burnout** compared to **Engagement** (OR = `{r} ORerr_burnout` , 95% CI\[`{r} lower_OR_burnout`,`{r} upper_OR_burnout`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_burnout`*,* *p* = `{r} p_values_burnout`.

**Having been diagnosed with Burnout** at **follow-up 2** was associated with 150% higher odds of being in **Risk** compared to **Engagement** (OR = `{r} ORerr_risk` , 95% CI\[`{r} lower_OR_risk`,`{r} upper_OR_risk`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_risk`*,* *p* = `{r} p_values_risk`.

### Transitions, pooled (3-profile) solution:

**Baseline -\> Follow-up 1:**

```{r}
#| label: Burnout predicting BL->F01 pooled

# Transition 1: Burnout0 → Burnout1 
fit_to_Burnout1 = glm(
  f01_class == "Burnout" ~ bl_class + bl_class:burnout_f01 - burnout_f01,
  family = binomial,
  data = lta_data_burnout)

summary(fit_to_Burnout1)

burnoutcoef = coef(summary(fit_to_Burnout1))
exp(burnoutcoef[4,1]) #1.14 (~14% increase in likelihood)

CI_low =  exp(burnoutcoef[4,1] - 1.96 * burnoutcoef[4,2]) #0.3
CI_high = exp(burnoutcoef[4,1] + 1.96 * burnoutcoef[4,2]) #3.8
p_val = burnoutcoef[4,4] #0.83

# Transition 2: Risk0 → Burnout1 
exp(burnoutcoef[5,1]) #2.68 (~168% increase in likelihood)

CI_low = exp(burnoutcoef[5,1] - 1.96 * burnoutcoef[5,2]) #1.1
CI_high = exp(burnoutcoef[5,1] + 1.96 * burnoutcoef[5,2]) #6.5
p_val = burnoutcoef[5,4] #0.03

# Transition 3: Engagement0 → Burnout1 
exp(burnoutcoef[6,1]) #0 (~0% increase in likelihood)

CI_low = exp(burnoutcoef[6,1] - 1.96 * burnoutcoef[6,2]) #0
CI_high = exp(burnoutcoef[6,1] + 1.96 * burnoutcoef[6,2]) #Inf
p_val = burnoutcoef[6,4] #0.99

# Transition 4: Burnout0 → Risk1 
fit_to_risk = glm(
  f01_class == "Risk" ~ bl_class + bl_class:burnout_f01 - burnout_f01,
  family = binomial,
  data = lta_data_burnout)

summary(fit_to_risk)

riskcoef = coef(summary(fit_to_risk))
exp(riskcoef[4,1]) #1.13 (~13% increase in likelihood)

CI_low = exp(riskcoef[4,1] - 1.96 * riskcoef[4,2]) #0.4
CI_high = exp(riskcoef[4,1] + 1.96 * riskcoef[4,2]) #3.6
p_val = riskcoef[4,4] #0.84

# Transition 5: Risk0 → Risk1 
exp(riskcoef[5,1]) #0.94 (~6% decrease in likelihood)

CI_low = exp(riskcoef[5,1] - 1.96 * riskcoef[5,2]) #0.5
CI_high = exp(riskcoef[5,1] + 1.96 * riskcoef[5,2]) #1.6
p_val = riskcoef[5,4] #0.81

# Transition 6: Engagement0 → Risk1 
exp(riskcoef[6,1]) #3.0 (~200% increase in likelihood)

CI_low = exp(riskcoef[6,1] - 1.96 * riskcoef[6,2]) #1.5
CI_high = exp(riskcoef[6,1] + 1.96 * riskcoef[6,2]) #5.8
p_val = riskcoef[6,4] #0.00

# Transition 7: Burnout0 → Engagement1
fit_to_engagement = glm(
  f01_class == "Engagement" ~ bl_class + bl_class:burnout_f01 - burnout_f01,
  family = binomial,
  data = lta_data_burnout)

summary(fit_to_engagement)
engagementcoef = coef(summary(fit_to_engagement))
exp(engagementcoef[4,1]) #0.44 (~56 % decrease in likelihood)

CI_low = exp(engagementcoef[4,1] - 1.96 * engagementcoef[4,2]) #0.04
CI_high = exp(engagementcoef[4,1] + 1.96 * engagementcoef[4,2]) #4.5
p_val = engagementcoef[4,4] #0.49

# Transition 8: Risk0 → Engagement1 
exp(engagementcoef[5,1]) #0.65 (~35% decrease in likelihood)

CI_low = exp(engagementcoef[5,1] - 1.96 * engagementcoef[5,2]) #0.3
CI_high = exp(engagementcoef[5,1] + 1.96 * engagementcoef[5,2]) #1.3
p_val = engagementcoef[5,4] #0.23

# Transition 9: Engagement0 → Engagement1 
exp(engagementcoef[6,1]) #0.34 (~66% decrease in likelihood)

CI_low = exp(engagementcoef[6,1] - 1.96 * engagementcoef[6,2]) #0.18
CI_high = exp(engagementcoef[6,1] + 1.96 * engagementcoef[6,2]) #0.66
p_val = engagementcoef[6,4] #0.00
```

**Follow-up 1 -\> Follow-up 2**:

```{r}
#| label: Burnout predicting F01->F02 pooled

# Transition 1: Burnout1 → Burnout2 
fit_to_Burnout1 = glm(
  f02_class == "Burnout" ~ f01_class + f01_class:burnout_f02 - burnout_f02,
  family = binomial,
  data = lta_data_burnout)

summary(fit_to_Burnout1)

burnoutcoef = coef(summary(fit_to_Burnout1))
exp(burnoutcoef[4,1]) #1.17 (~17% increase in likelihood)

CI_low =  exp(burnoutcoef[4,1] - 1.96 * burnoutcoef[4,2]) #0.3
CI_high = exp(burnoutcoef[4,1] + 1.96 * burnoutcoef[4,2]) #4.6
p_val = burnoutcoef[4,4] #0.82

# Transition 2: Risk1 → Burnout2 
exp(burnoutcoef[5,1]) #1.7 (~70% increase in likelihood)

CI_low = exp(burnoutcoef[5,1] - 1.96 * burnoutcoef[5,2]) #1.1
CI_high = exp(burnoutcoef[5,1] + 1.96 * burnoutcoef[5,2]) #2.6
p_val = burnoutcoef[5,4] #0.02

# Transition 3: Engagement1 → Burnout2 
exp(burnoutcoef[6,1]) #7.4 (~640% increase in likelihood)

CI_low = exp(burnoutcoef[6,1] - 1.96 * burnoutcoef[6,2]) #0.5
CI_high = exp(burnoutcoef[6,1] + 1.96 * burnoutcoef[6,2]) #121
p_val = burnoutcoef[6,4] #0.16

# Transition 4: Burnout1 → Risk2 
fit_to_risk = glm(
  f02_class == "Risk" ~ f01_class + f01_class:burnout_f02 - burnout_f02,
  family = binomial,
  data = lta_data_burnout)

summary(fit_to_risk)

riskcoef = coef(summary(fit_to_risk))
exp(riskcoef[4,1]) #0.88 (~12% decrease in likelihood)

CI_low = exp(riskcoef[4,1] - 1.96 * riskcoef[4,2]) #0.2
CI_high = exp(riskcoef[4,1] + 1.96 * riskcoef[4,2]) #4.2
p_val = riskcoef[4,4] #0.86

# Transition 5: Risk1 → Risk2 
exp(riskcoef[5,1]) #0.90 (~10% decrease in likelihood)

CI_low = exp(riskcoef[5,1] - 1.96 * riskcoef[5,2]) #0.6
CI_high = exp(riskcoef[5,1] + 1.96 * riskcoef[5,2]) #1.4
p_val = riskcoef[5,4] #0.64

# Transition 6: Engagement1 → Risk2 
exp(riskcoef[6,1]) #2.6 (~160% increase in likelihood)

CI_low = exp(riskcoef[6,1] - 1.96 * riskcoef[6,2]) #1.2
CI_high = exp(riskcoef[6,1] + 1.96 * riskcoef[6,2]) #5.9
p_val = riskcoef[6,4] #0.02

# Transition 7: Burnout1 → Engagement2
fit_to_engagement = glm(
  f02_class == "Engagement" ~ f01_class + f01_class:burnout_f02 - burnout_f02,
  family = binomial,
  data = lta_data_burnout)

summary(fit_to_engagement)
engagementcoef = coef(summary(fit_to_engagement))
exp(engagementcoef[4,1]) #0.89 (62% decrease in likelihood)

CI_low = exp(engagementcoef[4,1] - 1.96 * engagementcoef[4,2]) #0.1
CI_high = exp(engagementcoef[4,1] + 1.96 * engagementcoef[4,2]) #7.1
p_val = engagementcoef[4,4] #0.91

# Transition 8: Risk1 → Engagement2 
exp(engagementcoef[5,1]) #0.56 (~44% decrease in likelihood)

CI_low = exp(engagementcoef[5,1] - 1.96 * engagementcoef[5,2]) #0.3
CI_high = exp(engagementcoef[5,1] + 1.96 * engagementcoef[5,2]) #1.0
p_val = engagementcoef[5,4] #0.06

# Transition 9: Engagement1 → Engagement2 
exp(engagementcoef[6,1]) #0.34 (~66% decrease in likelihood)

CI_low = exp(engagementcoef[6,1] - 1.96 * engagementcoef[6,2]) #0.15
CI_high = exp(engagementcoef[6,1] + 1.96 * engagementcoef[6,2]) #0.78
p_val = engagementcoef[6,4] #0.01
```

## PHQ-9 as predictor

```{r warning = FALSE, message = FALSE}
#| echo: false

setwd("C:/Users/simon/Nextcloud2/Shared/LPA - Simon/04_longitudinal_analyses/data")
library(readr)
depression_bl = read_delim("export_phq9_bl_20240620.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
depression_follow1 = read_delim("export_phq9_fu01_20240620.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
depression_follow2 = read_delim("export_phq9_fu02_20240620.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)

depression_bl = depression_bl |> 
  rename(case = ProbCode) |> 
  mutate(case = str_remove(case, "-bl$")) |> 
  dplyr::select(-PHQ9_10)

any(is.na(depression_bl[, paste0("PHQ9_", 1:9)]))

depression_bl = depression_bl |> 
  group_by(case) |> 
  mutate(depression_score = sum(c_across(starts_with("PHQ9_")))) |> 
  ungroup()
depression_bl = depression_bl |> 
  dplyr::select(case, depression_score) |> 
  rename(depression_score_bl = depression_score)

depression_follow1 = depression_follow1 |> 
  rename(case = ProbCode) |> 
  mutate(case = str_remove(case, "-fu01$")) |> 
  dplyr::select(-PHQ9_10)

any(is.na(depression_follow1[, paste0("PHQ9_", 1:9)]))

depression_follow1 = depression_follow1 |> 
  group_by(case) |> 
  mutate(depression_score = sum(c_across(starts_with("PHQ9_")))) |> 
  ungroup()
depression_follow1 = depression_follow1 |> 
  dplyr::select(case, depression_score) |> 
  rename(depression_score_f01 = depression_score)

depression_follow2 = depression_follow2 |> 
  rename(case = ProbCode) |> 
  mutate(case = str_remove(case, "-fu02$")) |> 
  dplyr::select(-PHQ9_10)

any(is.na(depression_follow2[, paste0("PHQ9_", 1:9)]))

depression_follow2 = depression_follow2 |> 
  group_by(case) |> 
  mutate(depression_score = sum(c_across(starts_with("PHQ9_")))) |> 
  ungroup()
depression_follow2 = depression_follow2 |> 
  dplyr::select(case, depression_score) |> 
  rename(depression_score_f02 = depression_score)


depression_data = left_join(depression_bl, depression_follow1, by = "case")
depression_data = left_join(depression_data, depression_follow2, by = "case")

lta_data_depression = left_join(lta_data, depression_data, by = "case")
```

The extent of missing data is not critical, but noteworthy: `{r} colSums(is.na(lta_data_depression["depression_score_bl"]))` , `{r} colSums(is.na(lta_data_depression["depression_score_f01"]))` , `{r} colSums(is.na(lta_data_depression["depression_score_f02"]))`

**Descriptive differences in PHQ-9 depression scores** between respective profiles:

```{r}
table_data_bl = table(lta_data_depression$bl_class, lta_data_depression$depression_score_bl)
table_data_bl
chisq.test(table_data_bl)
mean_burnout = with(lta_data_depression, mean(depression_score_bl[bl_class == "Burnout"], na.rm = TRUE))
mean_risk = with(lta_data_depression, mean(depression_score_bl[bl_class == "Risk"], na.rm = TRUE))
mean_engagement = with(lta_data_depression, mean(depression_score_bl[bl_class == "Engagement"], na.rm = TRUE))
```

**At baseline,** mean PHQ-9 scores for the burnout class amount to `{r} mean_burnout` , whereas the risk class has a mean PHQ-9 score of `{r} mean_risk` and the engagement class of `{r} mean_engagement` . These differences are **highly significant**, as shown by the Chi-squared test above.

```{r}
table_data_f01 = table(lta_data_depression$f01_class, lta_data_depression$depression_score_f01)
table_data_f01
chisq.test(table_data_f01)
mean_burnout = with(lta_data_depression, mean(depression_score_f01[f01_class == "Burnout"], na.rm = TRUE))
mean_risk = with(lta_data_depression, mean(depression_score_f01[f01_class == "Risk"], na.rm = TRUE))
mean_engagement = with(lta_data_depression, mean(depression_score_f01[f01_class == "Engagement"], na.rm = TRUE))
```

**At follow-up 1,** mean PHQ-9 scores for the burnout class amount to `{r} mean_burnout` , whereas the risk class has a mean PHQ-9 score of `{r} mean_risk` and the engagement class of `{r} mean_engagement` . These differences are **highly significant**, as shown by the Chi-squared test above.

```{r}
table_data_f02 = table(lta_data_depression$f02_class, lta_data_depression$depression_score_f02)
table_data_f02
chisq.test(table_data_f02)
mean_burnout = with(lta_data_depression, mean(depression_score_f02[f02_class == "Burnout"], na.rm = TRUE))
mean_risk = with(lta_data_depression, mean(depression_score_f02[f02_class == "Risk"], na.rm = TRUE))
mean_engagement = with(lta_data_depression, mean(depression_score_f02[f02_class == "Engagement"], na.rm = TRUE))
```

**At follow-up 2**, mean PHQ-9 scores for the burnout class amount to `{r} mean_burnout` , whereas the risk class has a mean PHQ-9 score of `{r} mean_risk` and the engagement class of `{r} mean_engagement` . These differences are **highly significant**, as shown by the Chi-squared test above.

### Class memberships, pooled (3-profile) solution:

**PHQ-9 predicting class memberships at baseline:**

```{r}
#| label: Depression predicting class memberships

lta_data_depressionbl = lta_data_depression |> 
  select(bl_class, depression_score_bl)

#set reference level
lta_data_depressionbl$bl_class = relevel(as.factor(lta_data_depressionbl$bl_class), ref = "Engagement")

# Class as outcome, err as predictor, weights = posterior probabilities
mod_err = multinom(bl_class ~ depression_score_bl, data = lta_data_depressionbl, trace = FALSE)
summary(mod_err)
baseline = summary(mod_err)

# extract coefficients + SEs
coef_mat = baseline$coefficients
se_mat   = baseline$standard.errors

##################
#Burnout
# lower and upper CI on log-odds scale
lower_log_burnout = coef_mat[1,2] - 1.96 * se_mat[1,2]
upper_log_burnout = coef_mat[1,2] + 1.96 * se_mat[1,2]

# exponentiate: OR CI
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
ORerr_burnout = exp(coef_mat[1,2])

# Wald z-test
z_stats_burnout = coef_mat[1,2] / se_mat[1,2]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

####################
#Risk
# lower and upper CI on log-odds scale
lower_log_risk = coef_mat[2,2] - 1.96 * se_mat[2,2]
upper_log_risk = coef_mat[2,2] + 1.96 * se_mat[2,2]

# exponentiate: OR CI
lower_OR_risk = exp(lower_log_risk)
upper_OR_risk = exp(upper_log_risk)

# odds ratios
ORerr_risk = exp(coef_mat[2,2])

# Wald z-test
z_stats_risk = coef_mat[2,2] / se_mat[2,2]
p_values_risk = 2 * (1 - pnorm(abs(z_stats_risk)))
```

**Higher PHQ-9 depression scores** at **baseline** were associated with **63% higher odds** of being in **Burnout** compared to **Engagement** (OR = `{r} ORerr_burnout` , 95% CI\[`{r} lower_OR_burnout`,`{r} upper_OR_burnout`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_burnout`*,* *p* = `{r} p_values_burnout`.

**Higher PHQ-9 depression scores** at **baseline** were associated with **35% higher odds** of being in **Risk** compared to **Engagement** (OR = `{r} ORerr_risk` , 95% CI\[`{r} lower_OR_risk`,`{r} upper_OR_risk`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_risk`*,* *p* = `{r} p_values_risk`.

**PHQ-9 predicting class memberships at follow-up 1**:

```{r}
#| label: Depression predicting class memberships f01

lta_data_depressionf01 = lta_data_depression |> 
  select(f01_class, depression_score_f01)

#set reference level
lta_data_depressionf01$f01_class = relevel(as.factor(lta_data_depressionf01$f01_class), ref = "Engagement")

# Class as outcome, err as predictor, weights = posterior probabilities
mod_err = multinom(f01_class ~ depression_score_f01, data = lta_data_depressionf01, trace = FALSE)
summary(mod_err)
baseline = summary(mod_err)

# extract coefficients + SEs
coef_mat = baseline$coefficients
se_mat   = baseline$standard.errors

##################
#Burnout
# lower and upper CI on log-odds scale
lower_log_burnout = coef_mat[1,2] - 1.96 * se_mat[1,2]
upper_log_burnout = coef_mat[1,2] + 1.96 * se_mat[1,2]

# exponentiate: OR CI
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
ORerr_burnout = exp(coef_mat[1,2])

# Wald z-test
z_stats_burnout = coef_mat[1,2] / se_mat[1,2]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

####################
#Risk
# lower and upper CI on log-odds scale
lower_log_risk = coef_mat[2,2] - 1.96 * se_mat[2,2]
upper_log_risk = coef_mat[2,2] + 1.96 * se_mat[2,2]

# exponentiate: OR CI
lower_OR_risk = exp(lower_log_risk)
upper_OR_risk = exp(upper_log_risk)

# odds ratios
ORerr_risk = exp(coef_mat[2,2])

# Wald z-test
z_stats_risk = coef_mat[2,2] / se_mat[2,2]
p_values_risk = 2 * (1 - pnorm(abs(z_stats_risk)))
```

**Higher PHQ-9 depression scores** at **follow-up 1** were associated with **86% higher odds** of being in **Burnout** compared to **Engagement** (OR = `{r} ORerr_burnout` , 95% CI\[`{r} lower_OR_burnout`,`{r} upper_OR_burnout`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_burnout`*,* *p* = `{r} p_values_burnout`.

**Higher PHQ-9 depression scores** at **follow-up 1** were associated with **52% higher odds** of being in **Risk** compared to **Engagement** (OR = `{r} ORerr_risk` , 95% CI\[`{r} lower_OR_risk`,`{r} upper_OR_risk`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_risk`*,* *p* = `{r} p_values_risk`.

**PHQ-9 predicting class memberships at follow-up 2:**

```{r}
#| label: Depression predicting class memberships f02

lta_data_depressionf02 = lta_data_depression |> 
  select(f02_class, depression_score_f02)

#set reference level
lta_data_depressionf02$f02_class = relevel(as.factor(lta_data_depressionf02$f02_class), ref = "Engagement")

# Class as outcome, err as predictor, weights = posterior probabilities
mod_err = multinom(f02_class ~ depression_score_f02, data = lta_data_depressionf02, trace = FALSE)
summary(mod_err)
baseline = summary(mod_err)

# extract coefficients + SEs
coef_mat = baseline$coefficients
se_mat   = baseline$standard.errors

##################
#Burnout
# lower and upper CI on log-odds scale
lower_log_burnout = coef_mat[1,2] - 1.96 * se_mat[1,2]
upper_log_burnout = coef_mat[1,2] + 1.96 * se_mat[1,2]

# exponentiate: OR CI
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
ORerr_burnout = exp(coef_mat[1,2])

# Wald z-test
z_stats_burnout = coef_mat[1,2] / se_mat[1,2]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

####################
#Risk
# lower and upper CI on log-odds scale
lower_log_risk = coef_mat[2,2] - 1.96 * se_mat[2,2]
upper_log_risk = coef_mat[2,2] + 1.96 * se_mat[2,2]

# exponentiate: OR CI
lower_OR_risk = exp(lower_log_risk)
upper_OR_risk = exp(upper_log_risk)

# odds ratios
ORerr_risk = exp(coef_mat[2,2])

# Wald z-test
z_stats_risk = coef_mat[2,2] / se_mat[2,2]
p_values_risk = 2 * (1 - pnorm(abs(z_stats_risk)))
```

**Higher PHQ-9 depression scores** at **follow-up 2** were associated with **82% higher odds** of being in **Burnout** compared to **Engagement** (OR = `{r} ORerr_burnout` , 95% CI\[`{r} lower_OR_burnout`,`{r} upper_OR_burnout`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_burnout`*,* *p* = `{r} p_values_burnout`.

**Higher PHQ-9 depression scores** at **follow-up 2** were associated with **46% higher odds** of being in **Risk** compared to **Engagement** (OR = `{r} ORerr_risk` , 95% CI\[`{r} lower_OR_risk`,`{r} upper_OR_risk`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_risk`*,* *p* = `{r} p_values_risk`.

### Transitions, pooled (3-profile) solution:

**PHQ-9 predicting transitions from Baseline -\> Follow-up 1:**

```{r}
#| label: PHQ-9 predicting BL->F01 pooled

# Transition 1: Burnout0 → Burnout1 
fit_to_Burnout1 = glm(
  f01_class == "Burnout" ~ bl_class + bl_class:depression_score_f01 - depression_score_f01,
  family = binomial,
  data = lta_data_depression)

summary(fit_to_Burnout1)

burnoutcoef = coef(summary(fit_to_Burnout1))
exp(burnoutcoef[4,1]) #1.17 (~17% increase in likelihood, 390% with 10 units)

CI_low =  exp(burnoutcoef[4,1] - 1.96 * burnoutcoef[4,2]) #1.0
CI_high = exp(burnoutcoef[4,1] + 1.96 * burnoutcoef[4,2]) #1.3
p_val = burnoutcoef[4,4] #0.03

# Transition 2: Risk0 → Burnout1 
exp(burnoutcoef[5,1]) #1.2 (~20% increase in likelihood, 500% with 10 units)

CI_low = exp(burnoutcoef[5,1] - 1.96 * burnoutcoef[5,2]) #1.1
CI_high = exp(burnoutcoef[5,1] + 1.96 * burnoutcoef[5,2]) #1.3
p_val = burnoutcoef[5,4] #0.00

# Transition 3: Engagement0 → Burnout1 
exp(burnoutcoef[6,1]) #1.35 (~35% increase in likelihood, 1800% with 10 units)

CI_low = exp(burnoutcoef[6,1] - 1.96 * burnoutcoef[6,2]) #0.9
CI_high = exp(burnoutcoef[6,1] + 1.96 * burnoutcoef[6,2]) #1.8
p_val = burnoutcoef[6,4] #0.09

# Transition 4: Burnout0 → Risk1 
fit_to_risk = glm(
  f01_class == "Risk" ~ bl_class + bl_class:depression_score_f01 - depression_score_f01,
  family = binomial,
  data = lta_data_depression)

summary(fit_to_risk)

riskcoef = coef(summary(fit_to_risk))
exp(riskcoef[4,1]) #0.93 (~7% decrease in likelihood, 48% with 10 units)

CI_low = exp(riskcoef[4,1] - 1.96 * riskcoef[4,2]) #0.8
CI_high = exp(riskcoef[4,1] + 1.96 * riskcoef[4,2]) #1.1
p_val = riskcoef[4,4] #0.29

# Transition 5: Risk0 → Risk1 
exp(riskcoef[5,1]) #1.1 (~10% increase in likelihood, 186% with 10 units)

CI_low = exp(riskcoef[5,1] - 1.96 * riskcoef[5,2]) #1.0
CI_high = exp(riskcoef[5,1] + 1.96 * riskcoef[5,2]) #1.2
p_val = riskcoef[5,4] #0.001

# Transition 6: Engagement0 → Risk1 
exp(riskcoef[6,1]) #1.51 (~51% increase in likelihood, 5000% with 10 units)

CI_low = exp(riskcoef[6,1] - 1.96 * riskcoef[6,2]) #1.4
CI_high = exp(riskcoef[6,1] + 1.96 * riskcoef[6,2]) #1.7
p_val = riskcoef[6,4] #0.00

# Transition 7: Burnout0 → Engagement1
fit_to_engagement = glm(
  f01_class == "Engagement" ~ bl_class + bl_class:depression_score_f01 - depression_score_f01,
  family = binomial,
  data = lta_data_depression)

summary(fit_to_engagement)
engagementcoef = coef(summary(fit_to_engagement))
exp(engagementcoef[4,1]) #0.75 (~25% decrease in likelihood, 94% with 10 units) 

CI_low = exp(engagementcoef[4,1] - 1.96 * engagementcoef[4,2]) #0.6
CI_high = exp(engagementcoef[4,1] + 1.96 * engagementcoef[4,2]) #1.1
p_val = engagementcoef[4,4] #0.06

# Transition 8: Risk0 → Engagement1 
exp(engagementcoef[5,1]) #0.73 (~27% decrease in likelihood, 95% with 10 units)

CI_low = exp(engagementcoef[5,1] - 1.96 * engagementcoef[5,2]) #0.7
CI_high = exp(engagementcoef[5,1] + 1.96 * engagementcoef[5,2]) #0.8
p_val = engagementcoef[5,4] #0.00

# Transition 9: Engagement0 → Engagement1 
exp(engagementcoef[6,1]) #0.65 (~35% decrease in likelihood, 98% with 10 units)

CI_low = exp(engagementcoef[6,1] - 1.96 * engagementcoef[6,2]) #0.6
CI_high = exp(engagementcoef[6,1] + 1.96 * engagementcoef[6,2]) #0.7
p_val = engagementcoef[6,4] #0.00
```

**PHQ-9 predicting transitions from Follow-up 1 -\> Follow-up 2**:

```{r}
#| label: PHQ-9 predicting F01->F02 pooled

# Transition 1: Burnout0 → Burnout1 
fit_to_Burnout1 = glm(
  f02_class == "Burnout" ~ f01_class + f01_class:depression_score_f02 - depression_score_f02,
  family = binomial,
  data = lta_data_depression)

summary(fit_to_Burnout1)

burnoutcoef = coef(summary(fit_to_Burnout1))
exp(burnoutcoef[4,1]) #1.27 (~27% increase in likelihood, 987% with 10 units)

CI_low =  exp(burnoutcoef[4,1] - 1.96 * burnoutcoef[4,2]) #1.1
CI_high = exp(burnoutcoef[4,1] + 1.96 * burnoutcoef[4,2]) #1.5
p_val = burnoutcoef[4,4] #0.01

# Transition 2: Risk0 → Burnout1 
exp(burnoutcoef[5,1]) #1.28 (~28% increase in likelihood, 1067% with 10 units)

CI_low = exp(burnoutcoef[5,1] - 1.96 * burnoutcoef[5,2]) #1.2
CI_high = exp(burnoutcoef[5,1] + 1.96 * burnoutcoef[5,2]) #1.4
p_val = burnoutcoef[5,4] #0.00

# Transition 3: Engagement0 → Burnout1 
exp(burnoutcoef[6,1]) #1.35 (~35% increase in likelihood, 1900% with 10 units)

CI_low = exp(burnoutcoef[6,1] - 1.96 * burnoutcoef[6,2]) #1.1
CI_high = exp(burnoutcoef[6,1] + 1.96 * burnoutcoef[6,2]) #1.7
p_val = burnoutcoef[6,4] #0.02

# Transition 4: Burnout0 → Risk1 
fit_to_risk = glm(
  f02_class == "Risk" ~ f01_class + f01_class:depression_score_f02 - depression_score_f02,
  family = binomial,
  data = lta_data_depression)

summary(fit_to_risk)

riskcoef = coef(summary(fit_to_risk))
exp(riskcoef[4,1]) #0.95 (~5% decrease in likelihood, 41% with 10 units)

CI_low = exp(riskcoef[4,1] - 1.96 * riskcoef[4,2]) #0.8
CI_high = exp(riskcoef[4,1] + 1.96 * riskcoef[4,2]) #1.1
p_val = riskcoef[4,4] #0.53

# Transition 5: Risk0 → Risk1 
exp(riskcoef[5,1]) #0.94 (~5% decrease in likelihood, 44% with 10 units)

CI_low = exp(riskcoef[5,1] - 1.96 * riskcoef[5,2]) #0.90
CI_high = exp(riskcoef[5,1] + 1.96 * riskcoef[5,2]) #0.99
p_val = riskcoef[5,4] #0.01

# Transition 6: Engagement0 → Risk1 
exp(riskcoef[6,1]) #1.31 (~31% increase in likelihood, 1370% with 10 units)

CI_low = exp(riskcoef[6,1] - 1.96 * riskcoef[6,2]) #1.2
CI_high = exp(riskcoef[6,1] + 1.96 * riskcoef[6,2]) #1.5
p_val = riskcoef[6,4] #0.00

# Transition 7: Burnout0 → Engagement1
fit_to_engagement = glm(
  f02_class == "Engagement" ~ f01_class + f01_class:depression_score_f02 - depression_score_f02,
  family = binomial,
  data = lta_data_depression)

summary(fit_to_engagement)
engagementcoef = coef(summary(fit_to_engagement))
exp(engagementcoef[4,1]) #0.38 (~62% decrease in likelihood, 99% with 10 units) 

CI_low = exp(engagementcoef[4,1] - 1.96 * engagementcoef[4,2]) #0.1
CI_high = exp(engagementcoef[4,1] + 1.96 * engagementcoef[4,2]) #1.2
p_val = engagementcoef[4,4] #0.09

# Transition 8: Risk0 → Engagement1 
exp(engagementcoef[5,1]) #0.72 (~28% decrease in likelihood, 95% with 10 units)

CI_low = exp(engagementcoef[5,1] - 1.96 * engagementcoef[5,2]) #0.7
CI_high = exp(engagementcoef[5,1] + 1.96 * engagementcoef[5,2]) #0.8
p_val = engagementcoef[5,4] #0.00

# Transition 9: Engagement0 → Engagement1 
exp(engagementcoef[6,1]) #0.74 (~26% decrease in likelihood, 95% with 10 units)

CI_low = exp(engagementcoef[6,1] - 1.96 * engagementcoef[6,2]) #0.70
CI_high = exp(engagementcoef[6,1] + 1.96 * engagementcoef[6,2]) #0.8
p_val = engagementcoef[6,4] #0.00
```

## Social Support as predictor

```{r}
#| echo: false

setwd("C:/Users/simon/Nextcloud2/Shared/LPA - Simon/04_longitudinal_analyses/data")
social_follow1 = read_delim("export_fsozu_fu01_20241023.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)
social_follow2 = read_delim("export_fsozu_fu02_20240620.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)

social_follow1 = social_follow1 |> 
  rename(case = ProbCode) |> 
  mutate(case = str_remove(case, "-fu01$")) 

any(is.na(social_follow1[, paste0("F_SozU_", 1:14)]))

social_follow1 = social_follow1 |> 
  group_by(case) |> 
  mutate(socialsupport = sum(c_across(starts_with("F_SozU_")))) |> 
  ungroup()
social_follow1 = social_follow1 |> 
  dplyr::select(case, socialsupport) |> 
  rename(socialsupportf01 = socialsupport)

social_follow2 = social_follow2 |> 
  rename(case = ProbCode) |> 
  mutate(case = str_remove(case, "-fu02$")) 

any(is.na(social_follow2[, paste0("F_SozU_", 1:14)]))

social_follow2 = social_follow2 |> 
  group_by(case) |> 
  mutate(socialsupport = sum(c_across(starts_with("F_SozU_")))) |> 
  ungroup()
social_follow2 = social_follow2 |> 
  dplyr::select(case, socialsupport) |> 
  rename(socialsupportf02 = socialsupport)

social_data = left_join(social_follow1, social_follow2, by = "case")
lta_data_social = left_join(lta_data, social_data, by = "case")
```

**Descriptive differences in Social Support between respective profiles:**

```{r}
table_data_f01 = table(lta_data_social$f01_class, lta_data_social$socialsupportf01)
table_data_f01
chisq.test(table_data_f01)
mean_burnout = with(lta_data_social, mean(socialsupportf01[f01_class == "Burnout"], na.rm = TRUE))
mean_risk = with(lta_data_social, mean(socialsupportf01[f01_class == "Risk"], na.rm = TRUE))
mean_engagement = with(lta_data_social, mean(socialsupportf01[f01_class == "Engagement"], na.rm = TRUE))
```

**At follow-up 1,** mean social support scores for the burnout class amounted to `{r} mean_burnout` , whereas the risk class had a higher social support score of `{r} mean_risk` and the engagement class of `{r} mean_engagement` . These differences are **highly significant**, as shown by the Chi-squared test above.

```{r}
table_data_f02 = table(lta_data_social$f02_class, lta_data_social$socialsupportf02)
table_data_f02
chisq.test(table_data_f02)
mean_burnout = with(lta_data_social, mean(socialsupportf02[f02_class == "Burnout"], na.rm = TRUE))
mean_risk = with(lta_data_social, mean(socialsupportf02[f02_class == "Risk"], na.rm = TRUE))
mean_engagement = with(lta_data_social, mean(socialsupportf02[f02_class == "Engagement"], na.rm = TRUE))
```

**At follow-up 2,** mean social support scores for the burnout class amounted to `{r} mean_burnout` , whereas the risk class had a higher social support score of `{r} mean_risk` and the engagement class of `{r} mean_engagement` . These differences are **highly significant**, as shown by the Chi-squared test above.

### Class memberships, pooled (3-profile) solution:

**Social Support predicting class memberships at Follow-up 1:**

```{r}
#| label: Social support predicting class memberships

lta_data_socialf01 = lta_data_social |> 
  select(f01_class, socialsupportf01)

#set reference level
lta_data_socialf01$f01_class = relevel(as.factor(lta_data_socialf01$f01_class), ref = "Engagement")

# Class as outcome, err as predictor, weights = posterior probabilities
mod_err = multinom(f01_class ~ socialsupportf01, data = lta_data_socialf01, trace = FALSE)
summary(mod_err)
baseline = summary(mod_err)

# extract coefficients + SEs
coef_mat = baseline$coefficients
se_mat   = baseline$standard.errors

##################
#Burnout
# lower and upper CI on log-odds scale
lower_log_burnout = coef_mat[1,2] - 1.96 * se_mat[1,2]
upper_log_burnout = coef_mat[1,2] + 1.96 * se_mat[1,2]

# exponentiate: OR CI
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
ORerr_burnout = exp(coef_mat[1,2])

# Wald z-test
z_stats_burnout = coef_mat[1,2] / se_mat[1,2]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

####################
#Risk
# lower and upper CI on log-odds scale
lower_log_risk = coef_mat[2,2] - 1.96 * se_mat[2,2]
upper_log_risk = coef_mat[2,2] + 1.96 * se_mat[2,2]

# exponentiate: OR CI
lower_OR_risk = exp(lower_log_risk)
upper_OR_risk = exp(upper_log_risk)

# odds ratios
ORerr_risk = exp(coef_mat[2,2])

# Wald z-test
z_stats_risk = coef_mat[2,2] / se_mat[2,2]
p_values_risk = 2 * (1 - pnorm(abs(z_stats_risk)))
```

**Higher social support** at **follow-up 1** was associated with **12% lower odds** of being in **Burnout** compared to **Engagement** (OR = `{r} ORerr_burnout` , 95% CI\[`{r} lower_OR_burnout`,`{r} upper_OR_burnout`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_burnout`*,* *p* = `{r} p_values_burnout`.

**Higher social support** at **follow-up 1** were associated with **7% lower odds** of being in **Risk** compared to **Engagement** (OR = `{r} ORerr_risk` , 95% CI\[`{r} lower_OR_risk`,`{r} upper_OR_risk`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_risk`*,* *p* = `{r} p_values_risk`.

**Social support predicting class memberships at Follow-up 2:**

```{r}
#| label: Social support predicting class memberships f02

lta_data_socialf02 = lta_data_social |> 
  select(f02_class, socialsupportf02)

#set reference level
lta_data_socialf02$f02_class = relevel(as.factor(lta_data_socialf02$f02_class), ref = "Engagement")

# Class as outcome, err as predictor, weights = posterior probabilities
mod_err = multinom(f02_class ~ socialsupportf02, data = lta_data_socialf02, trace = FALSE)
summary(mod_err)
baseline = summary(mod_err)

# extract coefficients + SEs
coef_mat = baseline$coefficients
se_mat   = baseline$standard.errors

##################
#Burnout
# lower and upper CI on log-odds scale
lower_log_burnout = coef_mat[1,2] - 1.96 * se_mat[1,2]
upper_log_burnout = coef_mat[1,2] + 1.96 * se_mat[1,2]

# exponentiate: OR CI
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
ORerr_burnout = exp(coef_mat[1,2])

# Wald z-test
z_stats_burnout = coef_mat[1,2] / se_mat[1,2]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

####################
#Risk
# lower and upper CI on log-odds scale
lower_log_risk = coef_mat[2,2] - 1.96 * se_mat[2,2]
upper_log_risk = coef_mat[2,2] + 1.96 * se_mat[2,2]

# exponentiate: OR CI
lower_OR_risk = exp(lower_log_risk)
upper_OR_risk = exp(upper_log_risk)

# odds ratios
ORerr_risk = exp(coef_mat[2,2])

# Wald z-test
z_stats_risk = coef_mat[2,2] / se_mat[2,2]
p_values_risk = 2 * (1 - pnorm(abs(z_stats_risk)))
```

**Higher social support** at **follow-up 1** was associated with **10% lower odds** of being in **Burnout** compared to **Engagement** (OR = `{r} ORerr_burnout` , 95% CI\[`{r} lower_OR_burnout`,`{r} upper_OR_burnout`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_burnout`*,* *p* = `{r} p_values_burnout`.

**Higher social support** at **follow-up 1** were associated with **5% lower odds** of being in **Risk** compared to **Engagement** (OR = `{r} ORerr_risk` , 95% CI\[`{r} lower_OR_risk`,`{r} upper_OR_risk`\]. Wald z-tests further indicated that this difference in likelihood was **significant**, *z* = `{r} z_stats_risk`*,* *p* = `{r} p_values_risk`.

### Transitions, pooled (3-profile) solution:

**Social support predicting transitions from Follow-up 1 -\> Follow-up 2:**

```{r}
#| label: Social support predicting F01->F02 pooled

# Transition 1: Burnout0 → Burnout1 
fit_to_Burnout1 = glm(
  f02_class == "Burnout" ~ f01_class + f01_class:socialsupportf02 - socialsupportf02,
  family = binomial,
  data = lta_data_social)

summary(fit_to_Burnout1)

burnoutcoef = coef(summary(fit_to_Burnout1))
exp(burnoutcoef[4,1]) #0.90 (~10% decrease in likelihood, ~64% decrease 10 units)

CI_low =  exp(burnoutcoef[4,1] - 1.96 * burnoutcoef[4,2]) #0.8
CI_high = exp(burnoutcoef[4,1] + 1.96 * burnoutcoef[4,2]) #0.9
p_val = burnoutcoef[4,4] #0.01

# Transition 2: Risk0 → Burnout1 
exp(burnoutcoef[5,1]) #0.95 (~5% decrease in likelihood, 38% decrease 10 units)

CI_low = exp(burnoutcoef[5,1] - 1.96 * burnoutcoef[5,2]) #0.93
CI_high = exp(burnoutcoef[5,1] + 1.96 * burnoutcoef[5,2]) #0.97
p_val = burnoutcoef[5,4] #0.00

# Transition 3: Engagement0 → Burnout1 
exp(burnoutcoef[6,1]) #0.95 (~5% decrease in likelihood, 37% decrease 10 units)

CI_low = exp(burnoutcoef[6,1] - 1.96 * burnoutcoef[6,2]) #0.8
CI_high = exp(burnoutcoef[6,1] + 1.96 * burnoutcoef[6,2]) #1.1
p_val = burnoutcoef[6,4] #0.46

# Transition 4: Burnout0 → Risk1 
fit_to_risk = glm(
  f02_class == "Risk" ~ f01_class + f01_class:socialsupportf02 - socialsupportf02,
  family = binomial,
  data = lta_data_social)

summary(fit_to_risk)

riskcoef = coef(summary(fit_to_risk))
exp(riskcoef[4,1]) #1.06 (~6% increase in likelihood, 70% with 10 units)

CI_low = exp(riskcoef[4,1] - 1.96 * riskcoef[4,2]) #0.99
CI_high = exp(riskcoef[4,1] + 1.96 * riskcoef[4,2]) #1.1
p_val = riskcoef[4,4] #0.14

# Transition 5: Risk0 → Risk1 
exp(riskcoef[5,1]) #1.01 (~1% increase in likelihood, 20% with 10 units)

CI_low = exp(riskcoef[5,1] - 1.96 * riskcoef[5,2]) #0.99
CI_high = exp(riskcoef[5,1] + 1.96 * riskcoef[5,2]) #1.04
p_val = riskcoef[5,4] #0.06

# Transition 6: Engagement0 → Risk1 
exp(riskcoef[6,1]) #0.94 (~6% decrease in likelihood, 45% with 10 units)

CI_low = exp(riskcoef[6,1] - 1.96 * riskcoef[6,2]) #0.91
CI_high = exp(riskcoef[6,1] + 1.96 * riskcoef[6,2]) #0.94
p_val = riskcoef[6,4] #0.00

# Transition 7: Burnout0 → Engagement1
fit_to_engagement = glm(
  f02_class == "Engagement" ~ f01_class + f01_class:socialsupportf02 - socialsupportf02,
  family = binomial,
  data = lta_data_social)

summary(fit_to_engagement)
engagementcoef = coef(summary(fit_to_engagement))
exp(engagementcoef[4,1]) #1.17 (~17% increase in likelihood, 360% with 10 units) 

CI_low = exp(engagementcoef[4,1] - 1.96 * engagementcoef[4,2]) #0.99
CI_high = exp(engagementcoef[4,1] + 1.96 * engagementcoef[4,2]) #1.4
p_val = engagementcoef[4,4] #0.05

# Transition 8: Risk0 → Engagement1 
exp(engagementcoef[5,1]) #1.04 (~4% increase in likelihood, 50% with 10 units)

CI_low = exp(engagementcoef[5,1] - 1.96 * engagementcoef[5,2]) #1.0
CI_high = exp(engagementcoef[5,1] + 1.96 * engagementcoef[5,2]) #1.1
p_val = engagementcoef[5,4] #0.00

# Transition 9: Engagement0 → Engagement1 
exp(engagementcoef[6,1]) #1.06 (~6% increase in likelihood, 85% with 10 units)

CI_low = exp(engagementcoef[6,1] - 1.96 * engagementcoef[6,2]) #1.02
CI_high = exp(engagementcoef[6,1] + 1.96 * engagementcoef[6,2]) #1.1
p_val = engagementcoef[6,4] #0.00
```

### Social Support as protective factor

We aim to identify a potential buffering effect of social support on the negative effects of burnout/depression/ERR on class membership and transitions. So does higher social support attenuate the association between burnout/depression/ERR and worse states?

```{r}
#| label: buffering effects of social support

buffering = left_join(lta_data_social, lta_data_burnout, by = "case")
buffering = left_join(buffering, lta_data_eri3, by = "case")
buffering = left_join(buffering, lta_data_depression, by = "case")

buffering = buffering |> 
  select(case, bl_class.x, f01_class.x, f02_class.x, socialsupportf01, socialsupportf02, burnout_bl, burnout_f01, burnout_f02, errbl, err1, err2, depression_score_bl, depression_score_f01, depression_score_f02)

buffering = buffering |> 
  mutate(
    socialsupportf01_c = scale(socialsupportf01),
    err1_c = scale(err1),
    err2_c = scale(err1),
    depressionf01_c = scale(depression_score_f01),
    depressionf02_c = scale(depression_score_f02))
buffering = buffering |> 
  rename(bl_class = bl_class.x,
         f01_class = f01_class.x,
         f02_class = f02_class.x)

```

**Social support buffering the detrimental effects of burnout on class membership at Follow-up 1:**

```{r}
buffering$f01_class = relevel(as.factor(buffering$f01_class),ref = "Engagement")

mod_burn = multinom(f01_class ~ socialsupportf01_c*burnout_f01, data = buffering, trace = FALSE)
summary(mod_burn)
burn = summary(mod_burn)

coef_mat = burn$coefficients
se_mat   = burn$standard.errors

lower_log_burnout = coef_mat[1,4] - 1.96 * se_mat[1,4]
upper_log_burnout = coef_mat[1,4] + 1.96 * se_mat[1,4]
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
OR_burnout = exp(coef_mat[1,4])

z_stats_burnout = coef_mat[1,4] / se_mat[1,4]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

ggeffect(mod_burn, terms = c("socialsupportf01_c", "burnout_f01"))
plot(ggeffect(mod_burn, terms = c("socialsupportf01_c", "burnout_f01")), colors = c("firebrick4", "goldenrod2"), use_theme = TRUE)

```

A multinomial logistic regression revealed a **significant interaction between burnout diagnosis and social support predicting membership in the Burnout class** (OR = `{r} OR_burnout` , 95% CI\[`{r} lower_OR_burnout`,`{r} upper_OR_burnout`\], p = `{r} p_values_burnout`). As illustrated in the figure, individuals with a burnout diagnosis showed a substantially higher probability of being in the Burnout class at low levels of social support. However, **this effect decreased with increasing social support, indicating a buffering effect**. No significant interaction was observed for the Risk class.

**Social support buffering the detrimental effects of a bad ERR on class membership at Follow-up 1:**

```{r}
#convert ERR to factor variable <1 and >1
buffering = buffering |> 
  mutate(err1_c = as.numeric(err1_c),
         err1_c_factor = case_when(err1_c <= 0 ~ "< 0",
                                   err1_c >= 0 ~ "> 0"))

mod_burn = multinom(f01_class ~ socialsupportf01_c*err1_c_factor, data = buffering, trace = FALSE)
summary(mod_burn)
burn = summary(mod_burn)

coef_mat = burn$coefficients
se_mat   = burn$standard.errors

lower_log_burnout = coef_mat[1,4] - 1.96 * se_mat[1,4]
upper_log_burnout = coef_mat[1,4] + 1.96 * se_mat[1,4]
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
OR_burnout = exp(coef_mat[1,4])

z_stats_burnout = coef_mat[1,4] / se_mat[1,4]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

#convert ERR to factor variable <1 and >1
buffering = buffering |> 
  mutate(err1_c = as.numeric(err1_c),
         err1_c_factor = case_when(err1_c <= 0 ~ "< 0",
                                   err1_c >= 0 ~ "> 0"))

plot(ggeffect(mod_burn, terms = c("socialsupportf01_c", "err1_c_factor")),colors = c("navy", "coral2"), use_theme = TRUE)

```

A multinomial logistic regression revealed a non-significant interaction between the effort-reward ratio and social support predicting membership in the Burnout class (OR = `{r} OR_burnout` , 95% CI\[`{r} lower_OR_burnout`,`{r} upper_OR_burnout`\], p = `{r} p_values_burnout`). As illustrated in the figure, individuals with a burnout diagnosis showed a substantially higher probability of being in the Burnout class at low levels of social support. However, **this effect decreased with increasing social support, indicating a buffering effect**. No significant interaction was observed for the Risk class.

**Interaction between ERI and burnout at Baseline:**

```{r}
buffering$bl_class = relevel(as.factor(buffering$bl_class),ref = "Engagement")

mod_burn = multinom(bl_class ~ errbl_c*burnout_f01, data = buffering, trace = FALSE)
summary(mod_burn)
burn = summary(mod_burn)

coef_mat = burn$coefficients
se_mat   = burn$standard.errors

lower_log_burnout = coef_mat[1,4] - 1.96 * se_mat[1,4]
upper_log_burnout = coef_mat[1,4] + 1.96 * se_mat[1,4]
lower_OR_burnout = exp(lower_log_burnout)
upper_OR_burnout = exp(upper_log_burnout)

# odds ratios
OR_burnout = exp(coef_mat[1,4])

z_stats_burnout = coef_mat[1,4] / se_mat[1,4]
p_values_burnout = 2 * (1 - pnorm(abs(z_stats_burnout)))

ggeffect(mod_burn, terms = c("socialsupportf01_c", "burnout_f01"))
plot(ggeffect(mod_burn, terms = c("socialsupportf01_c", "burnout_f01")), colors = c("firebrick4", "goldenrod2"), use_theme = TRUE)
```

**Interaction between ERI and burnout at Follow-up 1:**

**Interaction between ERI and burnout at Follow-up 2:**

**Interaction between ERI and PHQ-9 at Baseline:**

**Interaction between ERI and PHQ-9 at Follow-up 1:**

**Interaction between ERI and PHQ-9 at Follow-up 2:**

**Social support buffering the detrimental effects of depression on class membership at Follow-up 1:**

Here I could even turn PHQ-9 scores into categories and then plot the results as well...

**Social support buffering the detrimental effects of burnout on class membership at Follow-up 2:**

**Social support buffering the detrimental effects of a bad ERR on class membership at Follow-up 2:**

**Social support buffering the detrimental effects of depression on class membership at Follow-up 2:**

**Social support buffering the detrimental effects of burnout on the transition to worse states Baseline-\> Follow-up 1:**

```{r}
model <- glm(
  transition_worse ~ depression * social_support_c,
  data = your_data,
  family = binomial
)
#Probe the interaction:

library(interactions)
interact_plot(
  mod_burn,
  pred = burnout_f01,		#or switch pred with modx?
  modx = socialsupportf01_c)
```
