---
title: "Latent profile analysis follow-up 1"
author: "Simon"
date: 2026-03-04
format: html
editor: visual
echo: true
bibliography: references.bib
---

## Preprocessing

```{r message=FALSE, warning=FALSE}
#| label: activate packages
library(readxl)
library(writexl)
library(car)
library(tidyverse)
library(tidyLPA)
library(mclust)
library(ggplot2)
library(ggalluvial)
library(forcats)
library(nnet)
```

We'll first assess latent burnout profiles in the entire samples (thus, without merging), which requires little data wrangling (datasets with prefix **full**).

```{r warning=FALSE, message=FALSE}
#| label: read data
setwd("~/EKFZ/02_coding/julian")
full_baseline = read_xlsx("mbi_final.xlsx")
full_follow1 = read.csv2("export_mbigs_fu01_20240620.csv")
full_follow2 = read.csv2("export_mbigs_fu02_20240620.csv")
```

```{r}
#| label: data wrangling

#remove suffixes
full_baseline = full_baseline |>  
  mutate(case = str_remove(case, "-bl$")) |> 
  group_by(case) |> 
  # reverse-code PE items  
  mutate(across(ends_with("_PE"),
      ~ 6 - .x, .names = "{.col}_IE")) 

full_follow1 = full_follow1 |> 
  rename(case = ProbCode) |> 
  mutate(case = str_remove(case, "-fu01$")) |> 
  group_by(case) |> 
  # reverse-code PE items  
  mutate(across(ends_with("_PE"),
      ~ 6 - .x, .names = "{.col}_IE")) 

full_follow2 = full_follow2 |> 
  rename(case = ProbCode) |> 
  mutate(case = str_remove(case, "-fu02$")) |> 
  group_by(case) |> 
  # reverse-code PE items  
  mutate(across(ends_with("_PE"),
      ~ 6 - .x, .names = "{.col}_IE"))
 
#select only relevant variables
full_baseline = full_baseline |> 
  dplyr::select(case, MBI1_EE, MBI2_EE, MBI3_EE, MBI4_EE, MBI5_PE_IE, MBI6_EE, MBI7_PE_IE,MBI8_CY, MBI9_CY, MBI10_PE_IE, MBI11_PE_IE, MBI12_PE, MBI13_CY, MBI14_CY, MBI15_CY, MBI16_PE_IE)

full_follow1 = full_follow1 |> 
  dplyr::select(case, MBI1_EE, MBI2_EE, MBI3_EE, MBI4_EE, MBI5_PE_IE, MBI6_EE, MBI7_PE_IE,MBI8_CY, MBI9_CY, MBI10_PE_IE, MBI11_PE_IE, MBI12_PE, MBI13_CY, MBI14_CY, MBI15_CY, MBI16_PE_IE)

full_follow2 = full_follow2 |> 
  dplyr::select(case, MBI1_EE, MBI2_EE, MBI3_EE, MBI4_EE, MBI5_PE_IE, MBI6_EE, MBI7_PE_IE,MBI8_CY, MBI9_CY, MBI10_PE_IE, MBI11_PE_IE, MBI12_PE, MBI13_CY, MBI14_CY, MBI15_CY, MBI16_PE_IE)
```

We'll then perform LPA on the reduced sample, including only those participants who enrolled for follow-up 2 (reducing the sample size to \~ 850, prefix **red**):

```{r}
#| label: add data
red_merged = right_join(full_baseline, full_follow1, by = "case")
red_merged = right_join(red_merged, full_follow2, by = "case")

red_baseline = red_merged |> 
  dplyr::select(case, MBI1_EE.x, MBI2_EE.x, MBI3_EE.x, MBI4_EE.x, MBI5_PE_IE.x, MBI6_EE.x, MBI7_PE_IE.x, MBI8_CY.x, MBI9_CY.x, MBI10_PE_IE.x, MBI11_PE_IE.x, MBI12_PE.x, MBI13_CY.x, MBI14_CY.x, MBI15_CY.x, MBI16_PE_IE.x)
red_baseline = na.omit(red_baseline)

red_follow1 = red_merged |> 
  dplyr::select(case, MBI1_EE.y, MBI2_EE.y, MBI3_EE.y, MBI4_EE.y, MBI5_PE_IE.y, MBI6_EE.y, MBI7_PE_IE.y, MBI8_CY.y, MBI9_CY.y, MBI10_PE_IE.y, MBI11_PE_IE.y, MBI12_PE.y, MBI13_CY.y, MBI14_CY.y, MBI15_CY.y, MBI16_PE_IE.y)
red_follow1 = na.omit(red_follow1)
red_follow1 = red_follow1[-c(840:1009), ]

red_follow2 = red_merged |> 
  dplyr::select(case, MBI1_EE, MBI2_EE, MBI3_EE, MBI4_EE, MBI5_PE_IE, MBI6_EE, MBI7_PE_IE,MBI8_CY, MBI9_CY, MBI10_PE_IE, MBI11_PE_IE, MBI12_PE, MBI13_CY, MBI14_CY, MBI15_CY, MBI16_PE_IE)
red_follow2 = na.omit(red_follow2)
red_follow2 = red_follow2[-c(840:1084), ]


```

Create mean scales for each of the timepoints so we can conduct **LPA** separately. Later, we will assess transitions and transition probabilities over time.

```{r}
#| label: mean scales for each sample

scales_full_bl = full_baseline |> 
  group_by(case) |> 
  mutate(
    exhaustion = if (any(is.na(c_across(ends_with("EE"))))) NA_real_
    else mean(c_across(ends_with("EE"))),
    inefficacy = if (any(is.na(c_across(ends_with("IE"))))) NA_real_
    else mean(c_across(ends_with("IE"))),
    cynicism = if (any(is.na(c_across(ends_with("CY"))))) NA_real_
    else mean(c_across(ends_with("CY")))) |> 
  ungroup() |> 
  dplyr::select(case, exhaustion, inefficacy, cynicism)
head(scales_full_bl)

scales_full_f1 = full_follow1|> 
  group_by(case) |> 
  mutate(
    exhaustion = if (any(is.na(c_across(ends_with("EE"))))) NA_real_
    else mean(c_across(ends_with("EE"))),
    inefficacy = if (any(is.na(c_across(ends_with("IE"))))) NA_real_
    else mean(c_across(ends_with("IE"))),
    cynicism = if (any(is.na(c_across(ends_with("CY"))))) NA_real_
    else mean(c_across(ends_with("CY")))) |> 
  ungroup() |> 
  dplyr::select(case, exhaustion, inefficacy, cynicism)
head(scales_full_f1)

scales_full_f2 = full_follow2 |> 
  group_by(case) |> 
  mutate(
    exhaustion = if (any(is.na(c_across(ends_with("EE"))))) NA_real_
    else mean(c_across(ends_with("EE"))),
    inefficacy = if (any(is.na(c_across(ends_with("IE"))))) NA_real_
    else mean(c_across(ends_with("IE"))),
    cynicism = if (any(is.na(c_across(ends_with("CY"))))) NA_real_
    else mean(c_across(ends_with("CY")))) |> 
  ungroup() |> 
  dplyr::select(case, exhaustion, inefficacy, cynicism)
head(scales_full_f2)

scales_red_bl = red_baseline |> 
  group_by(case) |> 
  mutate(
    exhaustion = if (any(is.na(c_across(ends_with("EE.x"))))) NA_real_
    else mean(c_across(ends_with("EE.x"))),
    inefficacy = if (any(is.na(c_across(ends_with("IE.x"))))) NA_real_
    else mean(c_across(ends_with("IE.x"))),
    cynicism = if (any(is.na(c_across(ends_with("CY.x"))))) NA_real_
    else mean(c_across(ends_with("CY.x")))) |> 
  ungroup() |> 
  dplyr::select(case, exhaustion, inefficacy, cynicism)
head(scales_red_bl)

scales_red_f1 = red_follow1|> 
  group_by(case) |> 
  mutate(
    exhaustion = if (any(is.na(c_across(ends_with("EE.y"))))) NA_real_
    else mean(c_across(ends_with("EE.y"))),
    inefficacy = if (any(is.na(c_across(ends_with("IE.y"))))) NA_real_
    else mean(c_across(ends_with("IE.y"))),
    cynicism = if (any(is.na(c_across(ends_with("CY.y"))))) NA_real_
    else mean(c_across(ends_with("CY.y")))) |> 
  ungroup() |> 
  dplyr::select(case, exhaustion, inefficacy, cynicism)
head(scales_red_f1)

scales_red_f2 = red_follow2 |> 
  group_by(case) |> 
  mutate(
    exhaustion = if (any(is.na(c_across(ends_with("EE"))))) NA_real_
    else mean(c_across(ends_with("EE"))),
    inefficacy = if (any(is.na(c_across(ends_with("IE"))))) NA_real_
    else mean(c_across(ends_with("IE"))),
    cynicism = if (any(is.na(c_across(ends_with("CY"))))) NA_real_
    else mean(c_across(ends_with("CY")))) |> 
  ungroup() |> 
  dplyr::select(case, exhaustion, inefficacy, cynicism)
head(scales_red_f2)

```

## LPA for entire sample, baseline:

```{r}
#| label: lpa entire sample, baseline


mbi_full_bl = scales_full_bl |> 
  dplyr::select(exhaustion, inefficacy, cynicism)

lpa = mbi_full_bl |>
  estimate_profiles(1:9,                        
                    variances = "equal",
                    covariances = "zero") 

#class enumeration with model fit indices
get_fit(lpa)
```

Evidently, the **5-class solution** seems to be superior:

1.  lower BIC than previous class solutions and marginal gap to the 6 class solution

2.  good, but lower entropy than other class solutions

3.  most robust sample size per class, prob_min and prob_max are solid

```{r}
#| label: decide class solution and extract class memberships full bl
lpa_5 = lpa[[5]]

mbi_profilesfullbl = lpa_5 |> 
  get_data()

# class models containing < 5% of the sample are rejected
c1 = with(mbi_profilesfullbl, length(which(Class == 1))) #2481
c2 = with(mbi_profilesfullbl, length(which(Class == 2))) #515
c3 = with(mbi_profilesfullbl, length(which(Class == 3))) #1288 
c4 = with(mbi_profilesfullbl, length(which(Class == 4))) #1458 
c5 = with(mbi_profilesfullbl, length(which(Class == 5))) #916 
```

Class 1 now contains `{r} c1` participants, class 2 contains `{r} c2` , class 3 `{r} c3` , class 4 `{r} c4` and class 5 `{r} c5`participants. Way to go! Now we will generate line plots that depict each classes' mean exhaustion, cynicism, and inefficacy scores:

```{r}
#| label: line plots full baseline

line_plotsfullbl = mbi_profilesfullbl |> 
  pivot_longer(
    cols = c(exhaustion, inefficacy, cynicism),
    names_to = "dimension",
    values_to = "score") |> 
  group_by(Class, dimension) |> 
  summarise(mean_score = mean(score),
            .groups = "drop")

#specify colors
profile_colorsfullbl = c(
  "2" = "midnightblue",
  "4" = "royalblue4",
  "3" = "royalblue3",
  "5" = "steelblue3",
  "1" = "lightskyblue")

line_plotsfullbl = line_plotsfullbl |> 
  mutate(Class = factor(Class)) |> 
  mutate(dimension = factor(dimension,levels = 
                              c("exhaustion", "cynicism", "inefficacy")))

fig1 = ggplot(line_plotsfullbl, aes(x = dimension, y = mean_score,
                              group = Class, color = Class)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = profile_colorsfullbl, breaks = c("2", "4", "3", "5", "1"), labels = c("Burnout","Disengaged","Overextended",
                       "Ineffective","Engagement")) +
  scale_x_discrete(labels = c(
    exhaustion = "Exhaustion",
    inefficacy   = "Inefficacy",
    cynicism   = "Cynicism")) +
  labs(y = "Mean Score", x = "", color = "Profile") +
   scale_y_continuous(limits = c(0, 6)) +
  theme_minimal()
plot(fig1)

#ggsave("full_bl.png", fig1, width = 12, height = 10, bg = "white", units = "cm", dpi = 400)
```

## LPA for entire sample, follow-up 1:

Within each profile, exhaustion, cynicism, inefficacy are assumed to be independent - any observed correlation is explained entirely by profile membership.

```{r warning = FALSE, message = FALSE}
#| label: lpa for timepoint full f1


mbi_full_f1 = scales_full_f1 |> 
  dplyr::select(exhaustion, inefficacy, cynicism)

lpa = mbi_full_f1 |>
  estimate_profiles(1:9,                        
                    variances = "equal",
                    covariances = "zero") 

#class enumeration with model fit indices
get_fit(lpa)
```

Evidently, the **5-class solution** seems to be superior:

1.  lower BIC than previous class solutions and marginal gap to the 4/5 class solution

2.  good, but lower entropy than other class solutions

3.  lowest ICL values

4.  most robust sample size per class, though 4% is slightly below the criterion

```{r}
#| label: decide class solution and extract class memberships full f1
lpa_5 = lpa[[5]]

mbi_profilesfullf1 = lpa_5 |> 
  get_data()

# class models containing < 5% of the sample are rejected
c1 = with(mbi_profilesfullf1, length(which(Class == 1))) #98 
c2 = with(mbi_profilesfullf1, length(which(Class == 2))) #427 
c3 = with(mbi_profilesfullf1, length(which(Class == 3))) #83 
c4 = with(mbi_profilesfullf1, length(which(Class == 4))) #709 
c5 = with(mbi_profilesfullf1, length(which(Class == 5))) #621 
```

Class 1 now contains `{r} c1` participants, class 2 contains `{r} c2` and class 3 `{r} c3`, class 4 `{r} c4` and class 5 `{r} c5`participants. Way to go! Now we will generate line plots that depict each classes' mean exhaustion, cynicism, and inefficacy scores:

```{r}
#| label: line plots full follow-up1

line_plotsfullf1 = mbi_profilesfullf1 |> 
  pivot_longer(
    cols = c(exhaustion, inefficacy, cynicism),
    names_to = "dimension",
    values_to = "score") |> 
  group_by(Class, dimension) |> 
  summarise(mean_score = mean(score),
            .groups = "drop")

#specify colors
profile_colorsfullf1 = c(
  "3" = "midnightblue",
  "2" = "royalblue4",
  "5" = "royalblue3",
  "1" = "steelblue3",
  "4" = "lightskyblue")

line_plotsfullf1 = line_plotsfullf1 |> 
  mutate(Class = factor(Class)) |> 
  mutate(dimension = factor(dimension,levels = 
                              c("exhaustion", "cynicism", "inefficacy")))

fig1 = ggplot(line_plotsfullf1, aes(x = dimension, y = mean_score,
                              group = Class, color = Class)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = profile_colorsfullf1, breaks = c("3", "2", "5", "1", "4"), labels = c("Burnout","Disengaged","Overextended",
                       "Ineffective","Engagement")) +
  scale_x_discrete(labels = c(
    exhaustion = "Exhaustion",
    inefficacy   = "Inefficacy",
    cynicism   = "Cynicism")) +
  labs(y = "Mean Score", x = "", color = "Profile") +
   scale_y_continuous(limits = c(0, 6)) +
  theme_minimal()
plot(fig1)

#ggsave("full_f1.png", fig1, width = 12, height = 10, bg = "white", units = "cm", dpi = 400)
```

## LPA for entire sample, follow-up 2:

```{r}
#| label: lpa for timepoint full f2


mbi_full_f2 = scales_full_f2 |> 
  dplyr::select(exhaustion, inefficacy, cynicism)

lpa = mbi_full_f2 |>
  estimate_profiles(1:9,                        
                    variances = "equal",
                    covariances = "zero") 

#class enumeration with model fit indices
get_fit(lpa)
```

Considering multiple fit indices, the **3-class solution** seems to be superior:

1.  lower BIC than previous class solutions

2.  most robust smallest class size of 18% - other class solutions fit better, but class sizes of \~1% are unacceptable.

```{r}
#| label: decide class solution and extract class memberships full f2

lpa_3 = lpa[[3]]
mbi_profilesfullf2 = lpa_3 |> 
  get_data()

# class models containing < 5% of the sample are rejected
c1 = with(mbi_profilesfullf2, length(which(Class == 1))) #356
c2 = with(mbi_profilesfullf2, length(which(Class == 2))) #197 
c3 = with(mbi_profilesfullf2, length(which(Class == 3))) #531 
```

Class 1 now contains `{r} c1` participants, class 2 contains `{r} c2` and class 3 `{r} c3` participants. Now we will generate line plots that depict each classes' mean exhaustion, cynicism, and inefficacy scores:

```{r}
#| label: line plots full follow-up2

line_plotsfullf2 = mbi_profilesfullf2 |> 
  pivot_longer(
    cols = c(exhaustion, inefficacy, cynicism),
    names_to = "dimension",
    values_to = "score") |> 
  group_by(Class, dimension) |> 
  summarise(mean_score = mean(score),
            .groups = "drop")

#specify colors
profile_colorsfullf2 = c(
  "2" = "midnightblue",
  "1" = "royalblue4",
  "3" = "royalblue3",
  "5" = "steelblue3",
  "4" = "lightskyblue")

line_plotsfullf2 = line_plotsfullf2 |> 
  mutate(Class = factor(Class)) |> 
  mutate(dimension = factor(dimension,levels = 
                              c("exhaustion", "cynicism", "inefficacy")))

fig1 = ggplot(line_plotsfullf2, aes(x = dimension, y = mean_score,
                              group = Class, color = Class)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = profile_colorsfullf2, breaks = c("2", "1", "3", "5", "4"), labels = c("Burnout","Disengaged","Engagement",
                       "Ineffective","Overextended")) +
  scale_x_discrete(labels = c(
    exhaustion = "Exhaustion",
    inefficacy   = "Inefficacy",
    cynicism   = "Cynicism")) +
  labs(y = "Mean Score", x = "", color = "Profile") +
   scale_y_continuous(limits = c(0, 6)) +
  theme_minimal()
plot(fig1)

#ggsave("full_f2.png", fig1, width = 12, height = 10, bg = "white", units = "cm", dpi = 400)
```

## LPA for reduced sample, baseline:

```{r}
#| label: lpa reduced sample, baseline


mbi_red_bl = scales_red_bl |> 
  dplyr::select(exhaustion, inefficacy, cynicism)

lpa = mbi_red_bl |>
  estimate_profiles(1:9,                        
                    variances = "equal",
                    covariances = "zero") 

#class enumeration with model fit indices
get_fit(lpa)
```

Evidently, the **5-class solution** seems to be superior:

1.  lowest BIC

2.  high entropy

3.  fairly small ICL

4.  n_max of 11% is better than previous and following class solutions

5.  prob_min and prob_max are promising

```{r}
#| label: decide class solution and extract class memberships reduced bl
lpa_5 = lpa[[5]]

mbi_profilesredbl = lpa_5 |> 
  get_data()

# class models containing < 5% of the sample are rejected
c1 = with(mbi_profilesredbl, length(which(Class == 1))) #206
c2 = with(mbi_profilesredbl, length(which(Class == 2))) #280
c3 = with(mbi_profilesredbl, length(which(Class == 3))) #106 
c4 = with(mbi_profilesredbl, length(which(Class == 4))) #91 
c5 = with(mbi_profilesredbl, length(which(Class == 5))) #156 
```

Class 1 now contains `{r} c1` participants, class 2 contains `{r} c2` , class 3 `{r} c3` , class 4 `{r} c4` and class 5 `{r} c5`participants. Way to go! Now we will generate line plots that depict each classes' mean exhaustion, cynicism, and inefficacy scores:

```{r}
#| label: line plots reduced baseline

line_plotsredbl = mbi_profilesredbl |> 
  pivot_longer(
    cols = c(exhaustion, inefficacy, cynicism),
    names_to = "dimension",
    values_to = "score") |> 
  group_by(Class, dimension) |> 
  summarise(mean_score = mean(score),
            .groups = "drop")

#specify colors
profile_colorsredbl = c(
  "3" = "midnightblue",
  "5" = "royalblue4",
  "1" = "royalblue3",
  "4" = "steelblue3",
  "2" = "lightskyblue")

line_plotsredbl = line_plotsredbl |> 
  mutate(Class = factor(Class)) |> 
  mutate(dimension = factor(dimension,levels = 
                              c("exhaustion", "cynicism", "inefficacy")))

fig1 = ggplot(line_plotsredbl, aes(x = dimension, y = mean_score,
                              group = Class, color = Class)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = profile_colorsredbl, breaks = c("3", "5", "1", "4", "2"), labels = c("Burnout","Disengaged","Overextended",
                       "Ineffective","Engagement")) +
  scale_x_discrete(labels = c(
    exhaustion = "Exhaustion",
    inefficacy   = "Inefficacy",
    cynicism   = "Cynicism")) +
  labs(y = "Mean Score", x = "", color = "Profile") +
   scale_y_continuous(limits = c(0, 6)) +
  theme_minimal()
plot(fig1)

#ggsave("red_bl.png", fig1, width = 12, height = 10, bg = "white", units = "cm", dpi = 400)
```

## LPA for reduced sample, follow-up 1:

```{r}
#| label: lpa for timepoint reduced f1


mbi_red_f1 = scales_red_f1 |> 
  dplyr::select(exhaustion, inefficacy, cynicism)

lpa = mbi_red_f1 |>
  estimate_profiles(1:9,                        
                    variances = "equal",
                    covariances = "zero") 

#class enumeration with model fit indices
get_fit(lpa)
```

Evidently, the **5-class solution** seems to be superior:

1.  lower BIC than previous class solutions and marginal gap to the 4/5 class solution

2.  good, but lower entropy than other class solutions

3.  lowest ICL values

4.  most robust sample size per class, though 4% is slightly below the criterion

```{r}
#| label: decide class solution and extract class memberships reduced f1
lpa_5 = lpa[[5]]

mbi_profilesredf1 = lpa_5 |> 
  get_data()

# class models containing < 5% of the sample are rejected
c1 = with(mbi_profilesredf1, length(which(Class == 1))) #177 
c2 = with(mbi_profilesredf1, length(which(Class == 2))) #46 
c3 = with(mbi_profilesredf1, length(which(Class == 3))) #321 
c4 = with(mbi_profilesredf1, length(which(Class == 4))) #30 
c5 = with(mbi_profilesredf1, length(which(Class == 5))) #265 
```

Class 1 now contains `{r} c1` participants, class 2 contains `{r} c2` and class 3 `{r} c3`, class 4 `{r} c4` and class 5 `{r} c5`participants. Way to go! Now we will generate line plots that depict each classes' mean exhaustion, cynicism, and inefficacy scores:

```{r}
#| label: line plots reduced follow-up1

line_plotsredf1 = mbi_profilesredf1 |> 
  pivot_longer(
    cols = c(exhaustion, inefficacy, cynicism),
    names_to = "dimension",
    values_to = "score") |> 
  group_by(Class, dimension) |> 
  summarise(mean_score = mean(score),
            .groups = "drop")

#specify colors
profile_colorsredf1 = c(
  "4" = "midnightblue",
  "1" = "royalblue4",
  "5" = "royalblue3",
  "2" = "steelblue3",
  "3" = "lightskyblue")

line_plotsredf1 = line_plotsredf1 |> 
  mutate(Class = factor(Class)) |> 
  mutate(dimension = factor(dimension,levels = 
                              c("exhaustion", "cynicism", "inefficacy")))

fig1 = ggplot(line_plotsredf1, aes(x = dimension, y = mean_score,
                              group = Class, color = Class)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = profile_colorsredf1, breaks = c("4", "1", "5", "2", "3"), labels = c("Burnout","Disengaged","Overextended",
                       "Ineffective","Engagement")) +
  scale_x_discrete(labels = c(
    exhaustion = "Exhaustion",
    inefficacy   = "Inefficacy",
    cynicism   = "Cynicism")) +
  labs(y = "Mean Score", x = "", color = "Profile") +
   scale_y_continuous(limits = c(0, 6)) +
  theme_minimal()
plot(fig1)

#ggsave("red_f1.png", fig1, width = 12, height = 10, bg = "white", units = "cm", dpi = 400)
```

## LPA for reduced sample, follow-up 2:

```{r warning= FALSE, message=FALSE}
#| label: lpa for timepoint reduced f2

mbi_red_f2 = scales_red_f2 |> 
  dplyr::select(exhaustion, inefficacy, cynicism)

lpa = mbi_red_f2 |>
  estimate_profiles(1:9,                        
                    variances = "equal",
                    covariances = "zero") 

#class enumeration with model fit indices
get_fit(lpa)
```

Considering multiple fit indices, the **3-class solution** seems to be superior:

1.  fairly good BIC, ICL, and Entropy values

2.  most robust smallest class size of 19% - other class solutions fit better, but class sizes of \~1% are unacceptable.

```{r}
#| label: decide class solution and extract class memberships reduced f2

lpa_3 = lpa[[3]]
mbi_profilesredf2 = lpa_3 |> 
  get_data()

# class models containing < 5% of the sample are rejected
c1 = with(mbi_profilesredf2, length(which(Class == 1))) #159
c2 = with(mbi_profilesredf2, length(which(Class == 2))) #303 
c3 = with(mbi_profilesredf2, length(which(Class == 3))) #377 
```

Class 1 now contains `{r} c1` participants, class 2 contains `{r} c2` and class 3 `{r} c3` participants. Now we will generate line plots that depict each classes' mean exhaustion, cynicism, and inefficacy scores:

```{r}
#| label: line plots reduced follow-up2

line_plotsredf2 = mbi_profilesredf2 |> 
  pivot_longer(
    cols = c(exhaustion, inefficacy, cynicism),
    names_to = "dimension",
    values_to = "score") |> 
  group_by(Class, dimension) |> 
  summarise(mean_score = mean(score),
            .groups = "drop")

#specify colors
profile_colorsredf2 = c(
  "1" = "midnightblue",
  "2" = "royalblue4",
  "3" = "royalblue3",
  "5" = "steelblue3",
  "4" = "lightskyblue")

line_plotsredf2 = line_plotsredf2 |> 
  mutate(Class = factor(Class)) |> 
  mutate(dimension = factor(dimension,levels = 
                              c("exhaustion", "cynicism", "inefficacy")))

fig1 = ggplot(line_plotsredf2, aes(x = dimension, y = mean_score,
                              group = Class, color = Class)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = profile_colorsredf2, breaks = c("1", "2", "3"), 
                     labels = c("Disengaged","Overextended", "Engagement")) +
  scale_x_discrete(labels = c(
    exhaustion = "Exhaustion",
    inefficacy   = "Inefficacy",
    cynicism   = "Cynicism")) +
  labs(y = "Mean Score", x = "", color = "Profile") +
   scale_y_continuous(limits = c(0, 6)) +
  theme_minimal()
plot(fig1)

ggsave("red_f2.png", fig1, width = 12, height = 10, bg = "white", units = "cm", dpi = 400)
```

## Alluvial plots and LTA baseline - follow-up 2

Alluvial plots showcase the transitions between classes over time, while LTA (or rather an LTA-like approach proposed by and adapted from [@lund2025] quantifies transition probabilities.

```{r}
#| label: prepare alluvial plots baseline - follow-up2

#merge baseline, f01 and f02 means data

mbi_merged = full_join(scales_red_bl, scales_red_f1, by = "case")
mbi_merged = full_join(mbi_merged, scales_red_f2, by = "case")

mbi_merged = mbi_merged |> 
  rename(exhaustion_bl = exhaustion.x, cynicism_bl = cynicism.x,
         inefficacy_bl = inefficacy.x, exhaustion_f01 = exhaustion.y,
         cynicism_f01 = cynicism.y, inefficacy_f01 = inefficacy.y,
         exhaustion_f02 = exhaustion, cynicism_f02 = cynicism,
         inefficacy_f02 = inefficacy)

#assign class memberships for bl, f01, and f02

mbi_profilesredbl = mbi_profilesredbl |> 
  mutate(artificialID = 1:839)
mbi_merged = mbi_merged |> 
  mutate(artificialID = 1:839)

mbi_merged = left_join(mbi_merged, mbi_profilesredbl, by = "artificialID")
mbi_merged = mbi_merged |> 
  rename(bl_class = Class) |> 
  dplyr::select(-model_number, -classes_number, -exhaustion, -cynicism, -inefficacy, 
         -CPROB1, -CPROB2, -CPROB3, -CPROB4, -CPROB5)

mbi_profilesredf1 = mbi_profilesredf1 |> 
  mutate(artificialID = 1:839)

mbi_merged = left_join(mbi_merged, mbi_profilesredf1, by = "artificialID")
mbi_merged = mbi_merged |> 
  rename(f01_class = Class) |> 
  dplyr::select(-model_number, -classes_number, -exhaustion, -cynicism, -inefficacy, 
         -CPROB1, -CPROB2, -CPROB3, -CPROB4, -CPROB5)

mbi_profilesredf2 = mbi_profilesredf2 |> 
  mutate(artificialID = 1:839)

mbi_merged = left_join(mbi_merged, mbi_profilesredf2, by = "artificialID")
mbi_merged = mbi_merged |> 
  rename(f02_class = Class) |> 
  dplyr::select(-model_number, -classes_number, -exhaustion, -cynicism, -inefficacy,          -CPROB1, -CPROB2, -CPROB3)

mbi_merged = mbi_merged |> 
  mutate(bl_class = factor(bl_class),
         f01_class = factor(f01_class),
         f02_class = factor(f02_class))

#reverse arbitrary class membership numbers:
mbi_merged = mbi_merged |> 
  mutate(f01_class = fct_recode(f01_class,
      "1" = "5",
      "2" = "3",
      "3" = "4",
      "4" = "2",
      "5" = "1"),f02_class = fct_recode(f02_class,
       "1" = "2",
       "2" = "3",
       "5" = "1"))

mbi_merged = mbi_merged |> 
  mutate(bl_class = fct_recode(bl_class,"Overextended" = "1","Engagement" = "2","Burnout" = "3", "Ineffective" = "4", "Disengaged" = "5"),
         f01_class = fct_recode(f01_class,"Overextended" = "1","Engagement" = "2","Burnout" = "3", "Ineffective" = "4", "Disengaged" = "5"),
         f02_class = fct_recode(f02_class,"Overextended" = "1","Engagement" = "2","Disengaged" = "5"))

mbi_merged_alluv = mbi_merged
head(mbi_merged_alluv)
```

```{r warning=FALSE, message=FALSE}
#| label: alluvial plots baseline - follow-up 2

mbi_merged_alluv = mbi_merged_alluv |> 
  mutate(bl_class = fct_relevel(bl_class, "Burnout", "Disengaged", "Overextended", "Ineffective", "Engagement"),
    f01_class = fct_relevel(f01_class, "Burnout", "Disengaged", "Overextended", "Ineffective", "Engagement"),
    f02_class = fct_relevel(f02_class, "Disengaged", "Overextended", "Engagement"))

all1 = ggplot(mbi_merged_alluv, aes(y = 1, axis1 = bl_class, axis2 = f01_class, axis3 = f02_class)) +
  geom_alluvium(aes(fill = bl_class), width = 1/12) +
  geom_stratum(aes(fill = after_stat(stratum)), width = 1/12, alpha = 0.8,
    color = "grey30") +
  geom_label(stat = "stratum", aes(label = after_stat(stratum)), color = "black") +
  scale_x_discrete(limits = c("Baseline", "Follow-up 1", "Follow-up 2"), expand = c(.05, .05)) +
  scale_fill_manual(
    name = "Burnout profiles",values = c(
      "Burnout"      = "midnightblue",
      "Disengaged"   = "royalblue4",
      "Overextended" = "royalblue3",
      "Ineffective"   = "steelblue3",
      "Engagement" = "lightskyblue"),
    na.translate = FALSE) +
  ggtitle("Longitudinal Stability of Latent Burnout Profiles") +
  theme_minimal()

all1
ggsave("alluvial.png", all1, width = 24, height = 16, bg = "white", units = "cm", dpi = 600)
```

## Latent Transition Analysis

Latent transition analysis in R is unusual, but feasible, as @lund2025 nicely show. We can skip the maximum likelihood imputation part (no missing data) and start off right with predicted transition tables:

```{r}
#| label: predicted transitions baseline - follow-up 2

lta_data = mbi_merged |> 
  select(case,bl_class, f01_class, f02_class)

# Probability of being in each F01 class, by BL class
# Fit multinomial model: F01 class predicted by BL class
trans_model = multinom(f01_class ~ bl_class - 1, data = lta_data, trace = FALSE)

# Get predicted probabilities for each BL class
pred_probs = predict(trans_model, 
                     newdata = data.frame(bl_class = c("Burnout","Disengaged","Overextended", "Ineffective", "Engagement")), type = "probs")

# Create transition probability matrix
transition_matrix = as.data.frame(pred_probs)
rownames(transition_matrix) = c("Burnout BL","Disengaged BL","Overextended BL", "Ineffective BL", "Engagement BL") 
colnames(transition_matrix) = c("Burnout F01","Disengaged F01","Overextended F01", "Ineffective F01", "Engagement F01") 

print("Transition Probability Matrix Follow-up 1:")
prop_df = print(round(transition_matrix, 3))
prop_df

#write_xlsx(prop_df, "transitions_bl_f01.xlsx")
```

```{r}
# Probability of being in each F01 class, by F01 class
# Fit multinomial model: F02 class predicted by F01 class
trans_model = multinom(f02_class ~ f01_class - 1, data = lta_data, trace = FALSE)

# Get predicted probabilities for each BL class
pred_probs = predict(trans_model, 
                     newdata = data.frame(f01_class = c("Burnout","Disengaged","Overextended", "Ineffective", "Engagement")), type = "probs")

# Create transition probability matrix
transition_matrix = as.data.frame(pred_probs)
rownames(transition_matrix) = c("Burnout F01","Disengaged F01","Overextended F01", "Ineffective F01", "Engagement F01") 
colnames(transition_matrix) = c("Disengaged F02","Overextended F02", "Engagement F02") 

print("Transition Probability Matrix Follow-up 2:")
prop_df = print(round(transition_matrix, 3))
prop_df

#write_xlsx(prop_df, "transitions_bl_f02.xlsx")
```

And now, finally, quantify the likelihood of specific class memberships in session 2 by session 1 class membership using binary logistic regressions:

```{r}
#| label: binary logistic regressions BL-Follow-up 1

# Probability of being in "Burnout" at F01
fit_burnout = glm(f01_class == "Burnout" ~ bl_class - 1, 
                  family = binomial,
                  data = lta_data)

coefs_burnout = coef(fit_burnout)
probs_burnout = exp(coefs_burnout) / (1 + exp(coefs_burnout))

print("Probability of Burnout at F01 by BL class:")
print(round(probs_burnout, 3))

# Probability of being in "Overextended" at session 2
fit_overextended = glm(f01_class == "Overextended" ~ bl_class - 1, 
                       family = binomial, 
                       data = lta_data)
coefs_overextended = coef(fit_overextended)
probs_overextended = exp(coefs_overextended) / (1 + exp(coefs_overextended))

print("Probability of Overextended at F01 by BL class:")
print(round(probs_overextended, 3))

# Probability of being in "Engagement" at F01

fit_engagement = glm(f01_class == "Engagement" ~ bl_class - 1,  # adjust labels
                     family = binomial, 
                     data = lta_data)
coefs_engagement = coef(fit_engagement)
probs_engagement = exp(coefs_engagement) / (1 + exp(coefs_engagement))

print("Probability of Engagement at F01 by BL class:")
print(round(probs_engagement, 3))

# Probability of being in "Disengaged" at session 2
fit_disengaged = glm(f01_class == "Disengaged" ~ bl_class - 1, 
                       family = binomial, 
                       data = lta_data)
coefs_disengaged = coef(fit_disengaged)
probs_disengaged = exp(coefs_disengaged) / (1 + exp(coefs_disengaged))

print("Probability of Disengaged at F01 by BL class:")
print(round(probs_disengaged, 3))

# Probability of being in "Ineffective" at F01

fit_ineffective = glm(f01_class == "Ineffective" ~ bl_class - 1,  # adjust labels
                     family = binomial, 
                     data = lta_data)
coefs_ineffective = coef(fit_ineffective)
probs_ineffective = exp(coefs_ineffective) / (1 + exp(coefs_ineffective))

print("Probability of Ineffective at F01 by BL class:")
print(round(probs_ineffective, 3))

```

```{r}
#| label: binary logistic regressions follow-up 1 - follow-up 2

# Probability of being in "Disengaged" at F02
fit_burnout = glm(f02_class == "Disengaged" ~ f01_class - 1, 
                  family = binomial,
                  data = lta_data)

coefs_burnout = coef(fit_burnout)
probs_burnout = exp(coefs_burnout) / (1 + exp(coefs_burnout))

print("Probability of Disengaged at F02 by F01 class:")
print(round(probs_burnout, 3))

# Probability of being in "Overextended" at F02
fit_overextended = glm(f02_class == "Overextended" ~ f01_class - 1, 
                       family = binomial, 
                       data = lta_data)
coefs_overextended = coef(fit_overextended)
probs_overextended = exp(coefs_overextended) / (1 + exp(coefs_overextended))

print("Probability of Overextended at F02 by F01 class:")
print(round(probs_overextended, 3))

# Probability of being in "Engagement" at F02

fit_engagement = glm(f02_class == "Engagement" ~ f01_class - 1,  # adjust labels
                     family = binomial, 
                     data = lta_data)
coefs_engagement = coef(fit_engagement)
probs_engagement = exp(coefs_engagement) / (1 + exp(coefs_engagement))

print("Probability of Engagement at F02 by F01 class:")
print(round(probs_engagement, 3))
```
